---
title: "Pericardial T Cells TCRM and Ctrl"
author: "Nadine"
date: "2026-04-16"
output: 
html_document:
  self_contained: no
  code_folding: show
  number_sections: no
  toc: yes
  toc_depth: 3
  toc_float: true
  collapsed: no
  smooth_scroll: yes
editor_options:
  chunk_output_type: inline
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(warning = FALSE, message = FALSE, dev = c("png", "pdf"))
seed <- 1234
```

# Load packages
```{r}
suppressPackageStartupMessages({
  library(Seurat)
  library(ggpubr)
  library(pheatmap)
  library(dplyr)
  library(tidyverse)
  library(clusterProfiler)
  library(org.Mm.eg.db)
  library(enrichplot)
  library(DOSE)
  library(grid)
  library(ggplot2)
  library(here)
  library (tibble)
  library(scater)
  library(SingleCellExperiment)
  library(matrixStats)
})
```

# Load seurat object
```{r}
# Load merged and cleaned object containing T cells
basedir<- here()
fileNam <- paste0(basedir, "/data/pericarditis_allmerged_Tcells_cleaned_seurat.rds")
seuratT <- readRDS(fileNam)
```

## DimPlot Clusters {.tabset}

### Resolution 0.1
```{r}
Idents(seuratT) <-  seuratT$RNA_snn_res.0.25
DimPlot(seuratT, reduction = "umap", label = T)
```

### Resolution 0.2
```{r}
Idents(seuratT) <-  seuratT$RNA_snn_res.0.4
DimPlot(seuratT, reduction = "umap", label = T)
```




# Marker genes per cluster
```{r, include = T, eval = T}
# Set identity to resolution 0.4
Idents(seuratT) <-  seuratT$RNA_snn_res.0.4

# Run marker genes
seuratT_markers_all <- FindAllMarkers(object = seuratT, assay ="RNA",
                                     only.pos = TRUE, min.pct = 0.25,
                                     logfc.threshold = 0.25,
                                     test.use = "wilcox")

# Save marker genes as text file
basedir<- here()
basedir <- paste0(basedir, "/docs/")
write.table(
  seuratT_markers_all,
  file = paste0(basedir, "pericarditis_allmerged_Tcells_cleaned_0.4_markerGenes.txt"),
  row.names = FALSE,
  quote = FALSE,
  sep = "\t"
)
```


# Assign labels to T cell clusters
```{r}
seuratT$TclusterLabel <- seuratT$RNA_snn_res.0.4

seuratT$TclusterLabel <- "TclusterLabel"
seuratT$TclusterLabel[which(seuratT$RNA_snn_res.0.4 %in% c("0"))] <- "CD4 Tscm"
seuratT$TclusterLabel[which(seuratT$RNA_snn_res.0.4 %in% c("1"))] <- "CD4 Tem"
seuratT$TclusterLabel[which(seuratT$RNA_snn_res.0.4 %in% c("2"))] <- "CD4 Teff"
seuratT$TclusterLabel[which(seuratT$RNA_snn_res.0.4 %in% c("3"))] <- "gdTcell-1"
seuratT$TclusterLabel[which(seuratT$RNA_snn_res.0.4 %in% c("4"))] <- "CD8 Tscm"
seuratT$TclusterLabel[which(seuratT$RNA_snn_res.0.4 %in% c("5"))] <- "gdTcell-2"
seuratT$TclusterLabel[which(seuratT$RNA_snn_res.0.4 %in% c("6"))] <- "NKTcell"
seuratT$TclusterLabel[which(seuratT$RNA_snn_res.0.4 %in% c("7"))] <- "CD8 Teff"
seuratT$TclusterLabel[which(seuratT$RNA_snn_res.0.4 %in% c("8"))] <- "Tregs"
seuratT$TclusterLabel[which(seuratT$RNA_snn_res.0.4 %in% c("9"))] <- "CD4 Tcm"


```


# Define color palette
```{r}
# Color palette TclusterLabel
colPal_T <- c("#CDCDC1",  "#8C8C82", "#A3B18A","#B5522A" ,  "#2F4F4F" , "#CD6600",  "#2B2B2B" ,"#8B3E2F"  , "#CDC673",  "#2C3E70")
names(colPal_T) <- c("CD4 Tscm", "CD4 Tem", "CD4 Teff", "gdTcell-1", "CD8 Tscm", "gdTcell-2", "NKTcell", "CD8 Teff", "Tregs", "CD4 Tcm")

# Color palette Phenotypes
colPal_Phenotype <- c("#8B0000", "#FCDACB", "#E6E6E6")
names(colPal_Phenotype) <- c( "iCMP", "inflamed", "healthy")
```

# Metrics
```{r}
table(seuratT$orig.ident)
table(seuratT$cond)
table(seuratT$Phenotype)
table(seuratT$TclusterLabel)
table(seuratT$TclusterLabel, seuratT$cond)
```



## DimPlot Clusters {.tabset}

### FBclusterLabel
```{r}
Idents(seuratT) <-  seuratT$TclusterLabel
DimPlot(seuratT, reduction = "umap", cols=colPal_T, group.by= "TclusterLabel")
```

### Condition split
```{r}
DimPlot(seuratT, reduction = "umap", cols=colPal_T, label = F, group.by= "TclusterLabel", split.by = "cond")
```



## Feature plot {.tabset}

### Cd3e
```{r}
FeaturePlot(object = seuratT, features = "ENSMUSG00000032093.Cd3e", cols = c("#C1CDCD", "#800000"), reduction = "umap", order = T)
```

### Cd4
```{r}
FeaturePlot(object = seuratT, features = "ENSMUSG00000023274.Cd4", cols = c("#C1CDCD", "#800000"), reduction = "umap", order = T)
```

### Cd8a
```{r}
FeaturePlot(object = seuratT, features = "ENSMUSG00000053977.Cd8a", cols = c("#C1CDCD", "#800000"), reduction = "umap", order = T)
```

### Tcf7
```{r}
FeaturePlot(object = seuratT, features = "ENSMUSG00000000782.Tcf7", cols = c("#C1CDCD", "#800000"), reduction = "umap", order = T)
```

### Id2
```{r}
FeaturePlot(object = seuratT, features = "ENSMUSG00000020644.Id2", cols = c("#C1CDCD", "#800000"), reduction = "umap", order = T)
```

### Ccr7
```{r}
FeaturePlot(object = seuratT, features = "ENSMUSG00000037944.Ccr7", cols = c("#C1CDCD", "#800000"), reduction = "umap", order = T)
```

### Sell
```{r}
FeaturePlot(object = seuratT, features = "ENSMUSG00000026581.Sell", cols = c("#C1CDCD", "#800000"), reduction = "umap", order = T)
```

### Cd44
```{r}
FeaturePlot(object = seuratT, features = "ENSMUSG00000005087.Cd44", cols = c("#C1CDCD", "#800000"), reduction = "umap", order = T)
```

### Fas
```{r}
FeaturePlot(object = seuratT, features = "ENSMUSG00000024778.Fas", cols = c("#C1CDCD", "#800000"), reduction = "umap", order = T)
```

### Pdcd1
```{r}
FeaturePlot(object = seuratT, features = "ENSMUSG00000026285.Pdcd1", cols = c("#C1CDCD", "#800000"), reduction = "umap", order = T)
```

### Ifng
```{r}
FeaturePlot(object = seuratT, features =  "ENSMUSG00000055170.Ifng", cols = c("#C1CDCD", "#800000"), reduction = "umap", order = T)
```

### Il17a
```{r}
FeaturePlot(object = seuratT, features = "ENSMUSG00000025929.Il17a", cols = c("#C1CDCD", "#800000"), reduction = "umap", order = T)
```

### Mki67
```{r}
FeaturePlot(object = seuratT, features = "ENSMUSG00000031004.Mki67", cols = c("#C1CDCD", "#800000"), reduction = "umap", order = T)
```

### Foxp3
```{r}
FeaturePlot(object = seuratT, features = "ENSMUSG00000039521.Foxp3", cols = c("#C1CDCD", "#800000"), reduction = "umap", order = T)
```

### Trgc1
```{r}
FeaturePlot(object = seuratT, features =  "ENSMUSG00000076749.Trgc1", cols = c("#C1CDCD", "#800000"), reduction = "umap", order = T)
```

### Zbtb16
```{r}
FeaturePlot(object = seuratT, features = "ENSMUSG00000066687.Zbtb16", cols = c("#C1CDCD", "#800000"), reduction = "umap", order = T)
```

### Gzma
```{r}
FeaturePlot(object = seuratT, features = "ENSMUSG00000023132.Gzma", cols = c("#C1CDCD", "#800000"), reduction = "umap", order = T)
```

### Il6ra
```{r}
FeaturePlot(object = seuratT, features = "ENSMUSG00000027947.Il6ra", cols = c("#C1CDCD", "#800000"), reduction = "umap", order = T)
```

### Gzma
```{r}
FeaturePlot(object = seuratT, features = "ENSMUSG00000021756.Il6st", cols = c("#C1CDCD", "#800000"), reduction = "umap", order = T)
```


# Marker genes per cluster
```{r, include = T, eval = T}
# Set identity to TclusterLabel
Idents(seuratT) <-  seuratT$TclusterLabel

# Run marker genes
seuratT_markers_all <- FindAllMarkers(object = seuratT, assay ="RNA",
                                     only.pos = TRUE, min.pct = 0.25,
                                     logfc.threshold = 0.25,
                                     test.use = "wilcox")

# Save marker genes as text file
basedir<- here()
basedir <- paste0(basedir, "/docs/")
write.table(
  seuratT_markers_all,
  file = paste0(basedir, "pericarditis_allmerged_Tcells_cleaned_0.4_cL_markerGenes.txt"),
  row.names = FALSE,
  quote = FALSE,
  sep = "\t"
)
```


```{r, include = F, eval = F}
# Load MarkerGenes
basedir<- here()
fileNam <- paste0(basedir, "/docs/pericarditis_allmerged_Tcells_cleaned_0.4_cL_markerGenes.txt")
seuratT_MarkerGenes <- read.table(fileNam, header = TRUE, sep = "\t", stringsAsFactors = FALSE)
```


## Dotplot of selected marker genes {.tabset}

### Dotplot of selected marker genes (unordered)
```{r, fig.height=6, fig.width=5}

Idents(seuratT) <- seuratT$TclusterLabel

#Marker genes as simple vector
selMarker <-selMarker <- c("Cd3e", "Cd4", "Cd8a", "Zbtb16", "Sell", "Cd44", "Id2", "Foxp3", "Tcf7", "Ccr7", "Nkg7", "S1pr1", "Lef1", "Trac", "Klrd1", "Ncr1", "Ccl5", "Eomes", "Trgc1", "Il23r", "Cd28")

#Assign a full gene name to the short geneID
genes <- data.frame(gene=rownames(seuratT)) %>% mutate(geneID=gsub(".*\\.", "", gene)) %>% 
  filter(geneID %in% as.character(selMarker)) %>% arrange(factor(geneID, levels = rev(selMarker)))

#Create dataframe of cL
clusterDat <- data.frame(cluster=levels(seuratT)) %>%
  mutate(label = levels(seuratT))


#Scale the gene across all clusters
DotPlot(
  object=seuratT,
  features = genes$gene,
  cols = c("#2C3E50", "#F1E6A1"),
  col.min = -2.5,
  col.max = 2.5,
  dot.min = 0,
  dot.scale = 6,
  idents = NULL,
  group.by = NULL,
  split.by = NULL,
  cluster.idents = T, #TRUE= automatic order, FALSE= ordered 
  scale = T,
  scale.by = "radius",
  scale.min = NA,
  scale.max = NA
) + coord_flip() + scale_x_discrete(breaks=genes$gene, labels=genes$geneID)  + scale_y_discrete(breaks=clusterDat$cluster, labels=clusterDat$label) + theme(axis.text.x = element_text(angle = 90))

```


### Dotplot of selected marker genes (ordered)
```{r, fig.height=6, fig.width=5}

# Set cluster order
ordVec <-  c("CD4 Tscm", "CD4 Tcm", "CD4 Tem", "CD4 Teff", "Tregs", "CD8 Tscm", "CD8 Teff", "gdTcell-1",  "gdTcell-2", "NKTcell")

# Set identities
Idents(seuratT) <- seuratT$TclusterLabel
Idents(seuratT) <- factor(Idents(seuratT), levels = ordVec)

# Marker genes as simple vector
selMarker <-selMarker <- c("Cd3e", "Cd4", "Cd8a", "Zbtb16", "Sell", "Cd44", "Id2", "Foxp3", "Tcf7", "Ccr7", "Nkg7", "S1pr1", "Lef1", "Trac", "Klrd1", "Ncr1", "Ccl5", "Eomes", "Trgc1", "Il23r", "Cd28")

# Assign a full gene name to the short geneID
genes <- data.frame(gene=rownames(seuratT)) %>% mutate(geneID=gsub(".*\\.", "", gene)) %>% 
  filter(geneID %in% as.character(selMarker)) %>% arrange(factor(geneID, levels = rev(selMarker)))

# Scale the gene across all clusters
DotPlot(
  object        = seuratT,
  features      = genes$gene,
  cols          = c("#2C3E50", "#F1E6A1"),
  col.min       = -2.5,
  col.max       =  2.5,
  dot.min       = 0,
  dot.scale     = 6,
  cluster.idents = TRUE,
  scale         = TRUE,
  scale.by      = "radius"
) +
  coord_flip() + scale_x_discrete(breaks = genes$gene,labels = genes$geneID) + scale_y_discrete(limits = ordVec) + theme(axis.text.x = element_text(angle = 90))

```
## Effector T cell signatures {.tabset}

### Th17 signature
```{r}

# Th17 gene signature
Th17_genes <- c("Il17a", "Il17f", "Rorc", "Stat3")

# Convert seuratT object to sce object
sceTh17 <- as.SingleCellExperiment(seuratT)

# Match genes in dataset
gene_map <- data.frame(
  geneID = rownames(sceTh17),
  gene = sub(".*\\.", "", rownames(sceTh17))
)

signGenes <- gene_map$geneID[gene_map$gene %in% Th17_genes]

# Module score (average expression)
expr <- assay(sceTh17, "logcounts")[signGenes, , drop = FALSE]


sceTh17$Th17_score <- colMeans(expr)

# Z-score the module
sceTh17$Th17_z <- as.numeric(scale(sceTh17$Th17_score))

lims <- quantile(sceTh17$Th17_z, c(0.01, 0.99), na.rm = TRUE)
sceTh17$Th17_z_clipped <- pmin(pmax(sceTh17$Th17_z, lims[1]), lims[2])

# Order positive cells on top
ord <- order(sceTh17$Th17_z_clipped)

# Plot umap
plotUMAP(
  sceTh17[, ord],
  color_by = "Th17_z_clipped",
  point_size = 0.5
)
```

### Th1 signature
```{r}
# Th1 gene signature
Th1_genes <- c("Ifng", "Tbx21", "Stat4", "Stat1")

# Convert seuratT object to sce object
sceTh1 <- as.SingleCellExperiment(seuratT)

# Match genes in dataset
gene_map <- data.frame(
  geneID = rownames(sceTh1),
  gene = sub(".*\\.", "", rownames(sceTh1))
)

signGenes <- gene_map$geneID[gene_map$gene %in% Th1_genes]

# Module score (average expression)
expr <- assay(sceTh1, "logcounts")[signGenes, , drop = FALSE]


sceTh1$Th1_score <- colMeans(expr)

# Z-score the module
sceTh1$Th1_z <- as.numeric(scale(sceTh1$Th1_score))

lims <- quantile(sceTh1$Th1_z, c(0.01, 0.99), na.rm = TRUE)
sceTh1$Th1_z_clipped <- pmin(pmax(sceTh1$Th1_z, lims[1]), lims[2])

# Order positive cells on top
ord <- order(sceTh1$Th1_z_clipped)

# Plot umap
plotUMAP(
  sceTh1[, ord],
  color_by = "Th1_z_clipped",
  point_size = 0.5
)
```


## Cluster abundances

### Conditions
```{r, , fig.height=3, fig.width=4}

# Plot abundance of clusters across Conditions
datList <- list()
for(con in unique(seuratT$cond)){
  seuratTSub <- subset(seuratT, cond==con)
  dat_con <- as.data.frame(table(seuratTSub$TclusterLabel)) %>%
  mutate(percent=Freq/ncol(seuratTSub)) %>% mutate(cond=con)
  datList[[con]] <- dat_con
}
dat_all <- do.call("rbind", datList)

# Set cluster order
ordY <-  c("CD4 Tscm", "CD4 Tcm", "CD4 Tem", "CD4 Teff", "Tregs", "CD8 Tscm", "CD8 Teff", "gdTcell-1",  "gdTcell-2", "NKTcell")
dat_all$Var1 <- factor(dat_all$Var1, levels = ordY)


# Barplot with Conditions on x axis and cluster colors
ggbarplot(dat_all, x= "cond", y= "percent", fill = "Var1", legend = "right", legend.titel = "cluster", ylab = "frequency")+ scale_fill_manual(values = colPal_T)

```



## Violin plots per cluster {.tabset}

### Il6
```{r}
genes <- data.frame(gene=rownames(seuratT))
# Gene expression across all cluster
Idents(seuratT) <- seuratT$TclusterLabel

colPal_T <- c("#CDCDC1",  "#8C8C82", "#A3B18A","#B5522A" ,  "#2F4F4F" , "#CD6600",  "#2B2B2B" ,"#8B3E2F"  , "#CDC673",  "#2C3E70")
names(colPal_T) <- c("CD4 Tscm", "CD4 Tem", "CD4 Teff", "gdTcell-1", "CD8 Tscm", "gdTcell-2", "NKTcell", "CD8 Teff", "Tregs", "CD4 Tcm")

# Set sample order
ordX <- c("CD4 Tscm", "CD4 Tcm", "CD4 Tem", "CD4 Teff", "Tregs", "CD8 Tscm", "CD8 Teff", "gdTcell-1",  "gdTcell-2", "NKTcell")

# Violin plot
VlnPlot(object = seuratT, features = "ENSMUSG00000025746.Il6", cols = colPal_T, alpha = 0.05) + scale_x_discrete(limits = ordX)
```

### Il6ra
```{r}
VlnPlot(object = seuratT, features = "ENSMUSG00000027947.Il6ra", cols = colPal_T, alpha = 0.05) + scale_x_discrete(limits = ordX)
```

### Il6st
```{r}
VlnPlot(object = seuratT, features = "ENSMUSG00000021756.Il6st", cols = colPal_T, alpha = 0.05) + scale_x_discrete(limits = ordX)
```


## Violin plots split by condition {.tabset}

### Il6
```{r}
VlnPlot(object = seuratT, features = "ENSMUSG00000025746.Il6", cols = c("#D3D3D3", "#B45B5C"), split.by = "cond", split.plot = TRUE, alpha = 0.1) + scale_x_discrete(limits = ordX)
```

### Il6ra
```{r}
VlnPlot(object = seuratT, features = "ENSMUSG00000027947.Il6ra", cols = c("#D3D3D3", "#B45B5C"), split.by = "cond", split.plot = TRUE, alpha = 0.1) + scale_x_discrete(limits = ordX)
```

### Il6st
```{r}
VlnPlot(object = seuratT, features = "ENSMUSG00000021756.Il6st", cols = c("#D3D3D3", "#B45B5C"), split.by = "cond", split.plot = TRUE, alpha = 0.1) + scale_x_discrete(limits = ordX)
```

# Load average heatmap function
```{r avg heatmap funct}

avgHeatmap <- function(seuratFB, selGenes, colVecIdent, colVecCond=NULL,
                       ordVec=NULL, gapVecR=NULL, gapVecC=NULL,cc=FALSE,
                       cr=FALSE, condCol=FALSE){
  
  selGenes <- selGenes$gene
  
  ## assay data
  clusterAssigned <- as.data.frame(Idents(seuratFB)) %>%
  dplyr::mutate(cell=rownames(.))
  colnames(clusterAssigned)[1] <- "ident"
  seuratFBDat <- GetAssayData(seuratFB)
  
  ## genes of interest
  genes <- data.frame(gene=rownames(seuratFB)) %>% 
    mutate(geneID=gsub("^.*\\.", "", gene)) %>% filter(geneID %in% selGenes)

  ## matrix with averaged cnts per ident
  logNormExpres <- as.data.frame(t(as.matrix(
    seuratFBDat[which(rownames(seuratFBDat) %in% genes$gene),])))
  logNormExpres <- logNormExpres %>% dplyr::mutate(cell=rownames(.)) %>%
    dplyr::left_join(.,clusterAssigned, by=c("cell")) %>%
    dplyr::select(-cell) %>% dplyr::group_by(ident) %>%
    dplyr::summarise_all(mean)
  logNormExpresMa <- logNormExpres %>% dplyr::select(-ident) %>% as.matrix()
  rownames(logNormExpresMa) <- logNormExpres$ident
  logNormExpresMa <- t(logNormExpresMa)
  rownames(logNormExpresMa) <- gsub("^.*?\\.","",rownames(logNormExpresMa))
  
  ## remove genes if they are all the same in all groups
  ind <- apply(logNormExpresMa, 1, sd) == 0
  logNormExpresMa <- logNormExpresMa[!ind,]
  genes <- genes[!ind,]

  ## color columns according to cluster
  annotation_col <- as.data.frame(gsub("(^.*?_)","",
                                       colnames(logNormExpresMa)))%>%
    dplyr::mutate(celltype=gsub("(_.*$)","",colnames(logNormExpresMa)))
  colnames(annotation_col)[1] <- "col1"
  annotation_col <- annotation_col %>%
    dplyr::mutate(cond = gsub("(^[0-9]_?)","",col1)) %>%
    dplyr::select(cond, celltype)
  rownames(annotation_col) <- colnames(logNormExpresMa) 

  ann_colors = list(
      #cond = colVecCond,
      celltype=colVecIdent)
  if(is.null(ann_colors$cond)){
    annotation_col$cond <- NULL
  }
  
  ## adjust order
  logNormExpresMa <- logNormExpresMa[selGenes,]
  if(is.null(ordVec)){
    ordVec <- levels(seuratFB)
  }
  logNormExpresMa <- logNormExpresMa[,ordVec]

  ## scaled row-wise
  pheatmap(logNormExpresMa, scale="row" ,treeheight_row = 0, cluster_rows = cr, 
         cluster_cols = cc,
         color = colorRampPalette(c("#2166AC", "#F7F7F7", "#B2182B"))(50),
         annotation_col = annotation_col, cellwidth=15, cellheight=10,
         annotation_colors = ann_colors, gaps_row = gapVecR, gaps_col = gapVecC)
}
```


### all T cells
```{r, fig.height=4, fig.width=5}

#genes <- data.frame(gene=rownames(seuratT))

selMarker <- list(stroma = c("Il17a", "Il17f", "Rorc", "Stat3", "Ifng", "Tbx21", "Stat4", "Stat1", "Foxp3"))
    
Idents(seuratT) <- seuratT$TclusterLabel

selGenes <- data.frame(gene=unlist(selMarker)) %>%
  rownames_to_column(var="grp") %>% mutate(Grp=gsub("_.*", "", grp))
grpCnt <- selGenes %>% group_by(Grp) %>% summarise(cnt=n())
gapR <- data.frame(Grp=unique(selGenes$Grp)) %>% 
  left_join(.,grpCnt, by="Grp") %>% mutate(cumSum=cumsum(cnt)) 


colPal_T <- c("#CDCDC1",  "#8C8C82", "#A3B18A","#B5522A" ,  "#2F4F4F" , "#CD6600",  "#2B2B2B" ,"#8B3E2F"  , "#CDC673",  "#2C3E70")
names(colPal_T) <- c("CD4 Tscm", "CD4 Tem", "CD4 Teff", "gdTcell-1", "CD8 Tscm", "gdTcell-2", "NKTcell", "CD8 Teff", "Tregs", "CD4 Tcm")

ordVec <- c("CD4 Tscm", "CD4 Tcm", "CD4 Tem", "CD4 Teff", "Tregs", "CD8 Tscm", "CD8 Teff", "gdTcell-1",  "gdTcell-2", "NKTcell")


pOut <- avgHeatmap(seurat = seuratT, selGenes = selGenes,
                  colVecIdent = colPal_T, 
                  ordVec=ordVec,
                  gapVecR=NULL, gapVecC=NULL,cc=F,
                  cr=T, condCol=F)
```

### ab T cells only
```{r, fig.height=4, fig.width=5}

#genes <- data.frame(gene=rownames(seuratT))

seuratTSub <- subset(seuratT, TclusterLabel %in% c("gdTcell-1", "gdTcell-2", "NKTcell"), invert=TRUE)

selMarker <- list(stroma = c("Il17a", "Il17f", "Rorc", "Stat3", "Ifng", "Tbx21", "Stat4", "Stat1", "Foxp3"))
    
Idents(seuratTSub) <- seuratTSub$TclusterLabel

selGenes <- data.frame(gene=unlist(selMarker)) %>%
  rownames_to_column(var="grp") %>% mutate(Grp=gsub("_.*", "", grp))
grpCnt <- selGenes %>% group_by(Grp) %>% summarise(cnt=n())
gapR <- data.frame(Grp=unique(selGenes$Grp)) %>% 
  left_join(.,grpCnt, by="Grp") %>% mutate(cumSum=cumsum(cnt)) 

colPal_T <- c("#CDCDC1",  "#8C8C82", "#A3B18A",  "#2F4F4F" ,  "#8B3E2F"  , "#CDC673",  "#2C3E70")
names(colPal_T) <- c("CD4 Tscm", "CD4 Tem", "CD4 Teff",  "CD8 Tscm",  "CD8 Teff", "Tregs", "CD4 Tcm")

ordVec <- c("CD4 Tscm", "CD4 Tcm", "CD4 Tem", "CD4 Teff", "Tregs", "CD8 Tscm", "CD8 Teff")


pOut <- avgHeatmap(seurat = seuratTSub, selGenes = selGenes,
                  colVecIdent = colPal_T, 
                  ordVec=ordVec,
                  gapVecR=NULL, gapVecC=NULL,cc=F,
                  cr=T, condCol=F)
```


# Save seurat object
```{r}
# Save seurat object
basedir <- here()
saveRDS(seuratT, file = paste0(basedir, "/data/pericarditis_allmerged_Tcells_cleaned_seurat_cL.rds"))
```


# Session info
```{r}
sessionInfo()
date()
```


