---
title: "characterize T cells"
author: "Nadine"
date: "2025-04-09"
output: 
html_document:
  self_contained: no
  code_folding: show
  number_sections: no
  toc: yes
  toc_depth: 3
  toc_float: true
  collapsed: no
  smooth_scroll: yes
editor_options:
  chunk_output_type: inline
---

```{r}
knitr::opts_chunk$set(warning = FALSE, message = FALSE, dev = c("png", "pdf"))
seed <- 1234
```

#load packages
```{r}
suppressPackageStartupMessages({
  library(ExploreSCdataSeurat3)
  library(runSeurat3)
  library(Seurat)
  library(ggpubr)
  library(pheatmap)
  library(SingleCellExperiment)
  library(dplyr)
  library(tidyverse)
  library(viridis)
  library(muscat)
  library(circlize)
  library(destiny)
  library(scater)
  library(metap)
  library(multtest)
  library(clusterProfiler)
  library(org.Mm.eg.db)
  library(msigdbr)
  library(enrichplot)
  library(DOSE)
  library(grid)
  library(gridExtra)
  library(ggupset)
  library(VennDiagram)
  library(reshape2)
  library(ggplot2)
  library(cowplot)
  library(purrr)
  library(here)
  library(ggsci)
  library(RColorBrewer)
})
```

#Load seurat object
```{r}

#Load merged and cleaned object from previous analysis
basedir<- here()
fileNam <- paste0(basedir, "/data/pericarditis_allmerged_Tcells_cleaned_seurat.rds")
seuratT <- readRDS(fileNam)

```


#Visualization in umap
```{r}
Idents(seuratT) <-  seuratT$RNA_snn_res.0.1
DimPlot(seuratT, reduction = "umap", group.by= "RNA_snn_res.0.1", label = T)

```

#Marker genes per cluster
```{r, include = T, eval = T}
#Set identity to resolution 0.1
Idents(seuratT) <-  seuratT$RNA_snn_res.0.1

#Run Marker Genes
seuratT_markers_all <- FindAllMarkers(object = seuratT, assay ="RNA",
                                     only.pos = TRUE, min.pct = 0.25,
                                     logfc.threshold = 0.25,
                                     test.use = "wilcox")

#Save marker genes as text file
basedir<- here()
basedir <- paste0(basedir, "/docs/")
write.table(
  seuratT_markers_all,
  file = paste0(basedir, "pericarditis_allmerged_Tcells_cleaned_0.1_markerGenes.txt"),
  row.names = FALSE,
  quote = FALSE,
  sep = "\t"
)

```

#Assign labels to Tcell clusters
```{r}
seuratT$TclusterLabel <- seuratT$RNA_snn_res.0.1

seuratT$TclusterLabel <- "TclusterLabel"
seuratT$TclusterLabel[which(seuratT$RNA_snn_res.0.1 %in% c("0"))] <- "CD4 Tscm"
seuratT$TclusterLabel[which(seuratT$RNA_snn_res.0.1 %in% c("1"))] <- "Tem"
seuratT$TclusterLabel[which(seuratT$RNA_snn_res.0.1 %in% c("2"))] <- "gd Tcell"
seuratT$TclusterLabel[which(seuratT$RNA_snn_res.0.1 %in% c("3"))] <- "CD8 Tscm"
seuratT$TclusterLabel[which(seuratT$RNA_snn_res.0.1 %in% c("4"))] <- "NKT"
```


#define cluster color palette
```{r}

colPal_T <- c("#CDCDC1","#CDC673", "#CD6600" ,"#8B3E2F","#2F4F4F" )
names(colPal_T) <- c("CD4 Tscm", "Tem", "gd Tcell", "NKT", "CD8 Tscm")

colPal_Phenotype <- c("#8B0000", "#FCDACB", "#E6E6E6")
names(colPal_Phenotype) <- c( "iCMP", "inflamed", "healthy")
```

#Stats
```{r}
table(seuratT$orig.ident)
table(seuratT$TclusterLabel)
table(seuratT$TclusterLabel, seuratT$cond)
table(seuratT$Phenotype)
table(seuratT$cond)
table(seuratT$TclusterLabel, seuratT$SampleName)
```

#DimPlot TclusterLabel
```{r}
Idents(seuratT) <-  seuratT$TclusterLabel

DimPlot(seuratT, reduction = "umap", cols=colPal_T, group.by= "TclusterLabel")
DimPlot(seuratT, reduction = "umap", cols=colPal_T, label = F, group.by= "TclusterLabel", split.by = "cond")
```

#DimPlot Condition
```{r}
DimPlot(seuratT, reduction = "umap", cols=c("#D3D3D3", "#B45B5C"), group.by = "cond")

DimPlot(seuratT, reduction = "umap", cols=c("#D3D3D3", "#B45B5C"), group.by = "cond", split.by = "cond")

```

#DimPlot Phenotype
```{r}
DimPlot(seuratT, reduction = "umap", cols=colPal_Phenotype, group.by = "Phenotype")

DimPlot(seuratT, reduction = "umap", group.by = "Phenotype", cols = colPal_Phenotype, order = c("iCMP", "inflamed","healthy"), pt.size = 0.02)


DimPlot(seuratT, reduction = "umap", cols=c("#E6E6E6","#E6E6E6","#FCDACB"), group.by = "Phenotype", order = "inflamed")
DimPlot(seuratT, reduction = "umap", cols=c("#E6E6E6","#E6E6E6","#8B0000"), group.by = "Phenotype", order = "iCMP")


```


#DimPlot Age
```{r}
DimPlot(seuratT, reduction = "umap", cols=c("#99B898", "#628395"), group.by = "age", split.by = "age")
DimPlot(seuratT, reduction = "umap", cols=c("#D3D3D3", "#628395"), group.by = "age", order = "8wk")
DimPlot(seuratT, reduction = "umap", cols=c("#D3D3D3", "#99B898"), group.by = "age", order = "4wk")
DimPlot(seuratT, reduction = "umap", cols=c("#D3D3D3", "#B45B5C"), group.by = "cond", split.by = "age")

```

# Feature plots - T cell identity
```{r}
#genes <- data.frame(gene=rownames(seuratT))


genes_sel <- c(
  "ENSMUSG00000032093.Cd3e",
  "ENSMUSG00000023274.Cd4",
  "ENSMUSG00000031004.Mki67",
  "ENSMUSG00000000782.Tcf7",
  "ENSMUSG00000037944.Ccr7",
  "ENSMUSG00000000409.Lck",
  "ENSMUSG00000026581.Sell",
  "ENSMUSG00000005087.Cd44",
  "ENSMUSG00000003882.Il7r",
  "ENSMUSG00000055170.Ifng",
  "ENSMUSG00000024778.Fas",
  "ENSMUSG00000020644.Id2",
  "ENSMUSG00000104876.Trdc",
  "ENSMUSG00000076749.Trgc1",
  "ENSMUSG00000076490.Trbc1",
  "ENSMUSG00000066687.Zbtb16",
  "ENSMUSG00000049093.Il23r",
  "ENSMUSG00000023132.Gzma",
  "ENSMUSG00000027947.Il6ra",
  "ENSMUSG00000025929.Il17a",
  "ENSMUSG00000021756.Il6st",
  "ENSMUSG00000053977.Cd8a"
)


plots <- lapply(genes_sel, function(gene) {
  FeaturePlot(
    object = seuratT,
    features = gene,
    cols = c("#C1CDCD", "#800000"),
    reduction = "umap", order = T
  )
})

for (p in plots) print(p)

```


#Marker genes per cluster
```{r, include = T, eval = T}
#Set identity to TclusterLabel
Idents(seuratT) <-  seuratT$TclusterLabel

#Run Marker Genes
seuratT_markers_all <- FindAllMarkers(object = seuratT, assay ="RNA",
                                     only.pos = TRUE, min.pct = 0.25,
                                     logfc.threshold = 0.25,
                                     test.use = "wilcox")

#Save marker genes as text file
basedir<- here()
basedir <- paste0(basedir, "/docs/")
write.table(
  seuratT_markers_all,
  file = paste0(basedir, "pericarditis_allmerged_Tcells_cleaned_0.1_cL_markerGenes.txt"),
  row.names = FALSE,
  quote = FALSE,
  sep = "\t"
)

```

#Load MarkerGenes
```{r}

basedir<- here()
fileNam <- paste0(basedir, "/docs/pericarditis_allmerged_Tcells_cleaned_0.1_cL_markerGenes.txt")
seuratT_MarkerGenes <- read.table(fileNam, header = TRUE, sep = "\t", stringsAsFactors = FALSE)
  
```


#Average heatmap function
```{r avg heatmap funct}

avgHeatmap <- function(seuratT, selGenes, colVecIdent, colVecCond=NULL,
                       ordVec=NULL, gapVecR=NULL, gapVecC=NULL,cc=FALSE,
                       cr=FALSE, condCol=FALSE){
  
  selGenes <- selGenes$gene
  
  ## assay data
  clusterAssigned <- as.data.frame(Idents(seuratT)) %>%
  dplyr::mutate(cell=rownames(.))
  colnames(clusterAssigned)[1] <- "ident"
  seuratTDat <- GetAssayData(seuratT)
  
  ## genes of interest
  genes <- data.frame(gene=rownames(seuratT)) %>% 
    mutate(geneID=gsub("^.*\\.", "", gene)) %>% filter(geneID %in% selGenes)

  ## matrix with averaged cnts per ident
  logNormExpres <- as.data.frame(t(as.matrix(
    seuratTDat[which(rownames(seuratTDat) %in% genes$gene),])))
  logNormExpres <- logNormExpres %>% dplyr::mutate(cell=rownames(.)) %>%
    dplyr::left_join(.,clusterAssigned, by=c("cell")) %>%
    dplyr::select(-cell) %>% dplyr::group_by(ident) %>%
    dplyr::summarise_all(mean)
  logNormExpresMa <- logNormExpres %>% dplyr::select(-ident) %>% as.matrix()
  rownames(logNormExpresMa) <- logNormExpres$ident
  logNormExpresMa <- t(logNormExpresMa)
  rownames(logNormExpresMa) <- gsub("^.*?\\.","",rownames(logNormExpresMa))
  
  ## remove genes if they are all the same in all groups
  ind <- apply(logNormExpresMa, 1, sd) == 0
  logNormExpresMa <- logNormExpresMa[!ind,]
  genes <- genes[!ind,]

  ## color columns according to cluster
  annotation_col <- as.data.frame(gsub("(^.*?_)","",
                                       colnames(logNormExpresMa)))%>%
    dplyr::mutate(celltype=gsub("(_.*$)","",colnames(logNormExpresMa)))
  colnames(annotation_col)[1] <- "col1"
  annotation_col <- annotation_col %>%
    dplyr::mutate(cond = gsub("(^[0-9]_?)","",col1)) %>%
    dplyr::select(cond, celltype)
  rownames(annotation_col) <- colnames(logNormExpresMa) 

  ann_colors = list(
      #cond = colVecCond,
      celltype=colVecIdent)
  if(is.null(ann_colors$cond)){
    annotation_col$cond <- NULL
  }
  
  ## adjust order
  logNormExpresMa <- logNormExpresMa[selGenes,]
  if(is.null(ordVec)){
    ordVec <- levels(seuratT)
  }
  logNormExpresMa <- logNormExpresMa[,ordVec]

  ## scaled row-wise
  pheatmap(logNormExpresMa, scale="row" ,treeheight_row = 0, cluster_rows = cr, 
         cluster_cols = cc,
         color = colorRampPalette(c("#2166AC", "#F7F7F7", "#B2182B"))(50),
         annotation_col = annotation_col, cellwidth=15, cellheight=10,
         annotation_colors = ann_colors, gaps_row = gapVecR, gaps_col = gapVecC)
}
```


#top 10 marker genes per cluster
```{r, fig.height=10}
cluster <- levels(seuratT)
selGenesAll <- seuratT_MarkerGenes %>% group_by(cluster) %>% 
  top_n(-10, p_val_adj) %>% 
  top_n(10, avg_log2FC)
names(selGenesAll) [names(selGenesAll) == "gene"] <- "Gene"
selGenesAll <- selGenesAll %>% mutate(gene=gsub("^.*\\.", "", Gene)) %>% filter(nchar(Gene)>1)

colPal_T <- c("#CDCDC1","#CDC673", "#CD6600" ,"#8B3E2F","#2F4F4F" )
names(colPal_T) <- c("CD4 Tscm", "Tem", "gd Tcell", "NKT", "CD8 Tscm")

ordVec <- names(colPal_T)

Idents(seuratT) <- seuratT$TclusterLabel
pOut <- avgHeatmap(seurat = seuratT, selGenes = selGenesAll,
                  colVecIdent = colPal_T, 
                  ordVec=ordVec,
                  gapVecR=NULL, gapVecC=NULL,cc=F,
                  cr=T, condCol=F)
```

#Dotplot of selected marker genes
```{r, fig.height=3.5, fig.width=4.5}

Idents(seuratT) <- seuratT$TclusterLabel

#Marker genes as simple vector
selMarker <- c("Cd3e", "Cd4", "Cd8a", "Zbtb16", "Sell", "Cd44", "Id2",  "Klrd1", "Foxp3", "Tcf7")

#Assign a full gene name to the short geneID
genes <- data.frame(gene=rownames(seuratT)) %>% mutate(geneID=gsub(".*\\.", "", gene)) %>% 
  filter(geneID %in% as.character(selMarker)) %>% arrange(factor(geneID, levels = rev(selMarker)))

#Create dataframe of cL
clusterDat <- data.frame(cluster=levels(seuratT)) %>%
  mutate(label = levels(seuratT))


#Scale the gene across all clusters
DotPlot(
  object=seuratT,
  features = genes$gene,
  cols = c("#2C3E50", "#F1E6A1"),
  col.min = -2.5,
  col.max = 2.5,
  dot.min = 0,
  dot.scale = 6,
  idents = NULL,
  group.by = NULL,
  split.by = NULL,
  cluster.idents = T, #TRUE= automatic order, FALSE= ordered 
  scale = T,
  scale.by = "radius",
  scale.min = NA,
  scale.max = NA
) + coord_flip() + scale_x_discrete(breaks=genes$gene, labels=genes$geneID)  + scale_y_discrete(breaks=clusterDat$cluster, labels=clusterDat$label) + theme(axis.text.x = element_text(angle = 90))

```



# T cell activation across conditions per cluster
```{r}
#Define cluster-condition label
seuratT$cluster_cond <- paste0(seuratT$TclusterLabel, "_", seuratT$cond)

#Define color palette for clusters
colPal_T <- c("#CDCDC1","#CDC673", "#CD6600" ,"#8B3E2F","#2F4F4F" )
names(colPal_T) <- c("CD4 Tscm", "Tem", "gd Tcell", "NKT", "CD8 Tscm")

#Set ordered cluster list
ordClusters <- names(colPal_T)

#Extract unique condition labels
ordConds <- unique(seuratT$cond)

#Order clusters by condition
ordVec_full <- as.vector(t(outer(ordClusters, ordConds, paste, sep = "_")))

#Assign as the factor levels for Seurat identities
Idents(seuratT) <- factor(seuratT$cluster_cond, levels = ordVec_full)

#Define marker genes
selMarker <- list(
  Tcells_ = c(
    "Mki67",
    "Tcf7", "Il2ra", "Ctla4",
    "Gata3", "Tbx21", "Rorc",
    "Ifng", "Il17a", "Sell",
    "Cd44", "Id2", "Prf1",
    "Foxp3"
  )
)

#Prepare selGenes and gapR
selGenes <- data.frame(gene = unlist(selMarker)) %>%
  tibble::rownames_to_column(var = "grp") %>%
  dplyr::mutate(Grp = gsub("_.*", "", grp))

grpCnt <- selGenes %>%
  dplyr::group_by(Grp) %>%
  dplyr::summarise(cnt = dplyr::n())

gapR <- data.frame(Grp = unique(selGenes$Grp)) %>%
  dplyr::left_join(grpCnt, by = "Grp") %>%
  dplyr::mutate(cumSum = cumsum(cnt))

#Generate heatmap
pOut <- avgHeatmap(
  seurat = seuratT,
  selGenes = selGenes,
  colVecIdent = colPal_T,
  ordVec = ordVec_full,
  gapVecR = gapR$cumSum,
  gapVecC = NULL,
  cc = FALSE,
  cr = TRUE,
  condCol = TRUE
)

```


#Abundance across non-sorted samples
```{r}
#Plot abundance of clusters in individual samples
##Leave out sorted samples

seuratTSub <- subset(seuratT, SampleName %in% c("8w_LMC8", "8w_TCRM7", "8w_TCRM8"), invert=TRUE )
  
##run calculations
datList <- NULL
for(con in unique(seuratTSub$SampleName)){
  seuratT_sub <- subset(seuratTSub, SampleName==con)
  print(dim(seuratT_sub))
  dat_SampleName <- as.data.frame(table(seuratT_sub$TclusterLabel)) %>%
  mutate(percent=Freq/ncol(seuratT_sub)) %>% mutate(SampleName=con)
  datList[[con]] <- dat_SampleName
}
dat_all <- do.call("rbind", datList)


ordX <- c("4w_LMC1", "4w_LMC2", "4w_LMC3", "4w_LMC4", "8w_LMC1", "8w_LMC2", "8w_LMC3", "8w_LMC4", "8w_LMC5", "8w_LMC6", "8w_LMC7",  "4w_TCRM1", "4w_TCRM2", "4w_TCRM3", "4w_TCRM4", "4w_TCRM5", "4w_TCRM6", "4w_TCRM7", "8w_TCRM1", "8w_TCRM2", "8w_TCRM3", "8w_TCRM4", "8w_TCRM5", "8w_TCRM6")

#Create barplot with samples on x axis and cluster colors
ggbarplot(dat_all, x= "SampleName", y= "percent", fill = "Var1", legend = "right", legend.titel = "cluster", ylab = "frequency") +  scale_fill_manual(values = colPal_T) + theme(axis.text.x = element_text(angle = 90)) + scale_x_discrete(limits=ordX) 



```

#Abundance across phenotypes
```{r, fig.height=3, fig.width=4}

datList <- NULL
for(con in unique(seuratT$Phenotype)){
  seuratTSub <- subset(seuratT, Phenotype==con)
  print(dim(seuratTSub))
  dat_Phenotype <- as.data.frame(table(seuratTSub$TclusterLabel)) %>%
  mutate(percent=Freq/ncol(seuratTSub)) %>% mutate(Phenotype=con)
  datList[[con]] <- dat_Phenotype
}
dat_all <- do.call("rbind", datList)

#Set order of Phenotypes
ordX <- c("healthy", "inflamed", "iCMP")

#Create barplot with Phenotypes on x axis and cluster colors
ggbarplot(dat_all, x= "Phenotype", y= "percent", fill = "Var1", legend = "right", legend.titel = "cluster", ylab = "frequency") +  scale_fill_manual(values = colPal_T) + theme(axis.text.x = element_text(angle = 90)) + scale_x_discrete(limits=ordX) 
```

#Abundance across conditions
```{r, , fig.height=3, fig.width=4}
datList <- NULL
for(con in unique(seuratT$cond)){
  seuratTSub <- subset(seuratT, cond==con)
  print(dim(seuratTSub))
  dat_con <- as.data.frame(table(seuratTSub$TclusterLabel)) %>%
  mutate(percent=Freq/ncol(seuratTSub)) %>% mutate(cond=con)
  datList[[con]] <- dat_con
}
dat_all <- do.call("rbind", datList)

## plot abundance
ggbarplot(dat_all, x= "cond", y= "percent", fill = "Var1", legend = "right", legend.titel = "cluster", ylab = "frequency")+ scale_fill_manual(values = colPal_T)

```


#Violin plots - Il6 receptors
```{r}
#genes <- data.frame(gene=rownames(seuratT))


genes_sel <- c(
  "ENSMUSG00000027947.Il6ra",
  "ENSMUSG00000021756.Il6st"
)

# gene expression in cluster
Idents(seuratT) <- seuratT$TclusterLabel

ordX <- names(colPal_T)

plots <- lapply(genes_sel, function(gene) {
  VlnPlot(
    object = seuratT,
    features = gene,
    cols = colPal_T, alpha = 0.05
  ) + scale_x_discrete(limits=ordX)
})

for (p in plots) print(p)



plots <- lapply(genes_sel, function(gene) {
  VlnPlot(
    object = seuratT,
    features = gene,
    cols = c("#D3D3D3", "#B45B5C"), split.by = "cond", split.plot = T, alpha = 0.1
  ) + scale_x_discrete(limits=ordX)
})

for (p in plots) print(p)


```


#Save seurat object
```{r}
#Save seurat object
basedir <- here()
saveRDS(seuratT, file = paste0(basedir, "/data/pericarditis_allmerged_Tcells_cleaned_seurat_cL.rds"))
```


#session info
```{r}
sessionInfo()
date()
```


