---
title: "12_cellchat TCRM vs Ctrl"
author: "Nadine"
date: "2025-07-14"
output: 
html_document:
  self_contained: no
  code_folding: show
  number_sections: no
  toc: yes
  toc_depth: 3
  toc_float: true
  collapsed: no
  smooth_scroll: yes
editor_options:
  chunk_output_type: inline
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(warning = FALSE, message = FALSE, dev = c("png", "pdf"))
seed <- 1234
```


#load packages
```{r}
suppressPackageStartupMessages({
  library(ExploreSCdataSeurat3)
  library(runSeurat3)
  library(Seurat)
  library(ggpubr)
  library(pheatmap)
  library(SingleCellExperiment)
  library(dplyr)
  library(tidyverse)
  library(viridis)
  library(muscat)
  library(circlize)
  library(destiny)
  library(scater)
  library(metap)
  library(multtest)
  library(clusterProfiler)
  library(org.Mm.eg.db)
  library(msigdbr)
  library(enrichplot)
  library(DOSE)
  library(grid)
  library(gridExtra)
  library(ggupset)
  library(VennDiagram)
  library(reshape2)
  library(ggplot2)
  library(cowplot)
  library(purrr)
  library(here)
  library(ggsci)
  library(RColorBrewer)
  library(CellChat)
  library(patchwork)
  library(NMF)
  library(circlize)
  library(ComplexHeatmap)
  library(BiocGenerics)
  library(BiocNeighbors)
  library(SeuratObject)
})


```


#load seurat object
```{r}
#Load full dataset seurat object from previous analysis
basedir<- here()
fileNam <- paste0(basedir, "/data/pericarditis_allmerged_cleaned_cL_seurat.rds")
seuratM <- readRDS(fileNam)

```


#Set color palette for clusters
```{r}
colPal <- c( "#CDC1C5","#8B7B8B","#CDC673","#CD5B45","#C1CDC1","#7AC5CD","#7F7F7F","#4A708B","#CD919E","#8B668B","#8B475D","#8FBC8F","#CCCCCC","#8B7D6B","#CDC8B1","#333333")

names(colPal) <- c("B2 Bcells","B1 Bcells", "Cycling cells","Plasma cells", "Tcells", "ILCs", "gdTcells",  "NKcells", "Mono/Mphs", "Neutrophils", "DCs", "Fibroblasts-1", "Fibroblasts-2", "Cover cells","Mesothelial cells", "Adipocytes") 

colPal_Phenotype <- c("#8B0000", "#CD4F39", "#FCDACB", "#E6E6E6")
names(colPal_Phenotype) <- c("DCM", "iCMP", "inflamed", "healthy")

```


#Set order of clusterLabel
```{r}
order <- c("B2 Bcells","B1 Bcells", "Cycling cells","Plasma cells", "Tcells", "ILCs", "gdTcells",  "NKcells", "Mono/Mphs", "Neutrophils", "DCs", "Fibroblasts-1", "Fibroblasts-2", "Cover cells","Mesothelial cells", "Adipocytes") 

seuratM$clusterLabel <- factor(seuratM$clusterLabel, levels = order)
```

#Run cellchat for LMC
```{r, include = T, eval = F}
#Subset for Controls
seurat_LMC <- subset(seuratM, cond %in% "LMC")
sce_LMC <- as.SingleCellExperiment(seurat_LMC)
rownames(sce_LMC) = gsub("^.*\\.","",rownames(sce_LMC))

#Create cellchat object
cellchat_LMC <- createCellChat(object = sce_LMC, group.by = "clusterLabel")

#Select database
CellChatDB <- CellChatDB.mouse
CellChatDB.use <- CellChatDB
cellchat_LMC@DB <- CellChatDB.use

#Run analysis
cellchat_LMC <- subsetData(cellchat_LMC)
cellchat_LMC <- identifyOverExpressedGenes(cellchat_LMC)
cellchat_LMC <- identifyOverExpressedInteractions(cellchat_LMC)

#Filter
options(future.globals.maxSize = 8000 * 1024^2) #necessary if limit too small
cellchat_LMC <- filterCommunication(cellchat_LMC, min.cells = 5)
cellchat_LMC <- computeCommunProb(cellchat_LMC, type =  "truncatedMean", trim = 0.2, population.size = FALSE)

cellchat_LMC <- computeCommunProbPathway(cellchat_LMC)

cellchat_LMC <- aggregateNet(cellchat_LMC)

future.seed=TRUE
cellchat_LMC <- netAnalysis_computeCentrality(cellchat_LMC, slot.name = "netP") 

```

#Save cellchat LMC
```{r, include = T, eval = F}
basedir <- here()
saveRDS(cellchat_LMC, file = paste0(basedir, "/data/pericarditis_allmerged_cleaned_cellchat_LMC.rds"))
```

#Run cellchat for TCRM
```{r, include = T, eval = F}
#Subset for Controls
seurat_TCRM <- subset(seuratM, cond %in% "TCRM")
sce_TCRM <- as.SingleCellExperiment(seurat_TCRM)
rownames(sce_TCRM) = gsub("^.*\\.","",rownames(sce_TCRM))

#Create cellchat object
cellchat_TCRM <- createCellChat(object = sce_TCRM, group.by = "clusterLabel")

#Select database
CellChatDB <- CellChatDB.mouse
CellChatDB.use <- CellChatDB
cellchat_TCRM@DB <- CellChatDB.use

#Run analysis
cellchat_TCRM <- subsetData(cellchat_TCRM)
cellchat_TCRM <- identifyOverExpressedGenes(cellchat_TCRM)
cellchat_TCRM <- identifyOverExpressedInteractions(cellchat_TCRM)

#Filter
options(future.globals.maxSize = 8000 * 1024^2) #necessary if limit too small
cellchat_TCRM <- filterCommunication(cellchat_TCRM, min.cells = 5)
cellchat_TCRM <- computeCommunProb(cellchat_TCRM, type =  "truncatedMean", trim = 0.2, population.size = FALSE)

cellchat_TCRM <- computeCommunProbPathway(cellchat_TCRM)

cellchat_TCRM <- aggregateNet(cellchat_TCRM)

future.seed=TRUE
cellchat_TCRM <- netAnalysis_computeCentrality(cellchat_TCRM, slot.name = "netP") 

```

#Save cellchat TCRM
```{r, include = T, eval = F}
basedir <- here()
saveRDS(cellchat_TCRM, file = paste0(basedir, "/data/pericarditis_allmerged_cleaned_cellchat_TCRM.rds"))
```



#Load cellchat files
```{r}

basedir<- here()
fileNam <- paste0(basedir, "/data/pericarditis_allmerged_cleaned_cellchat_LMC.rds")
cellchat_LMC <- readRDS(fileNam)


basedir<- here()
fileNam <- paste0(basedir, "/data/pericarditis_allmerged_cleaned_cellchat_TCRM.rds")
cellchat_TCRM <- readRDS(fileNam)


```


# Interaction analysis of IL6 across conditions
## LMC
### Calculate the aggregated cell-cell communication network
```{r}
#Signaling role analysis on the aggregated cell-cell communication network from all signaling pathways
gg1 <- netAnalysis_signalingRole_scatter(cellchat_LMC, color.use = colPal)
gg1

```


### Heatmap of sender/ receiver (selected pathways)
```{r}
#cellchat_all@netP$pathway

# Define pathways to analyze
pathways.show <- c("IL6")

# Compute centrality
cellchat_LMC <- netAnalysis_computeCentrality(cellchat_LMC, slot.name = "netP")

# Loop through each pathway and plot the network role
for (pathway in pathways.show) {
  netAnalysis_signalingRole_network(
    cellchat_LMC,
    signaling = pathway,
    width = 8,
    height = 2.5,
    font.size = 10,
    color.use = colPal,
    color.heatmap = "Greys"
  )
}
```

### Chord diagram for selected pathways
```{r}


# Define signaling pathways you to visualize
pathways.show <- c("IL6")


# Loop through each and generate the chord diagram
for (pathway in pathways.show) {
  netVisual_aggregate(
    cellchat_LMC,
    signaling = pathway,
    layout = "chord",
    color.use = colPal,
    title.name = pathway
  )
}


```


##TCRM
### Calculate the aggregated cell-cell communication network
```{r}
# Signaling role analysis on the aggregated cell-cell communication network from all signaling pathways
gg1 <- netAnalysis_signalingRole_scatter(cellchat_TCRM, color.use = colPal)
gg1

```

### Heatmap of sender/ receiver (selected pathways)
```{r}
#cellchat_all@netP$pathway

# Define pathways to analyze
pathways.show <- c("IL6")

# Compute centrality
cellchat_TCRM <- netAnalysis_computeCentrality(cellchat_TCRM, slot.name = "netP")

# Loop through each pathway and plot the network role
for (pathway in pathways.show) {
  netAnalysis_signalingRole_network(
    cellchat_TCRM,
    signaling = pathway,
    width = 8,
    height = 2.5,
    font.size = 10,
    color.use = colPal,
    color.heatmap = "Reds"
  )
}
```

### Chord diagram for selected pathways
```{r}


# Define signaling pathways you to visualize
pathways.show <- c("IL6")


# Loop through each and generate the chord diagram
for (pathway in pathways.show) {
  netVisual_aggregate(
    cellchat_TCRM,
    signaling = pathway,
    layout = "chord",
    color.use = colPal,
    title.name = pathway
  )
}


```


