---
title: "CellChat Pericardial Cells"
author: "Nadine"
date: "2026-04-17"
output: 
html_document:
  self_contained: no
  code_folding: show
  number_sections: no
  toc: yes
  toc_depth: 3
  toc_float: true
  collapsed: no
  smooth_scroll: yes
editor_options:
  chunk_output_type: inline
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(warning = FALSE, message = FALSE, dev = c("png", "pdf"))
seed <- 1234
```


# Load packages
```{r}
suppressPackageStartupMessages({
  library(Seurat)
  library(SingleCellExperiment)
  library(here)
  library(CellChat)
})
```


# Load seurat object
```{r}
# Load merged and cleaned object containing FBs
basedir<- here()
fileNam <- paste0(basedir, "/data/pericarditis_allmerged_cleaned_cL_seurat.rds")
seuratM <- readRDS(fileNam)
```


# Define color palette
```{r}
# Color palette clusterLabel
colPal <- c( "#CDC1C5","#8B7B8B","#CDC673","#CD5B45","#C1CDC1","#7AC5CD","#7F7F7F","#4A708B","#CD919E","#8B668B","#8B475D","#8FBC8F","#CCCCCC","#8B7D6B","#CDC8B1","#333333")

names(colPal) <- c("B2 Bcells","B1 Bcells", "Cycling cells","Plasma cells", "Tcells", "ILCs", "gdTcells",  "NKcells", "Mono/Mphs", "Neutrophils", "DCs", "Fibroblasts-1", "Fibroblasts-2", "Cover cells","Mesothelial cells", "Adipocytes") 

```


# Set order of clusterLabel
```{r}
order <- c("B2 Bcells","B1 Bcells", "Cycling cells","Plasma cells", "Tcells", "ILCs", "gdTcells",  "NKcells", "Mono/Mphs", "Neutrophils", "DCs", "Fibroblasts-1", "Fibroblasts-2", "Cover cells","Mesothelial cells", "Adipocytes") 

seuratM$clusterLabel <- factor(seuratM$clusterLabel, levels = order)
```


# Run CellChat for Ctrl
```{r, include = T, eval = F}
# Subset Controls
seurat_LMC <- subset(seuratM, cond %in% "LMC")
sce_LMC <- as.SingleCellExperiment(seurat_LMC)
rownames(sce_LMC) = gsub("^.*\\.","",rownames(sce_LMC))

# Create CellChat object
cellchat_LMC <- createCellChat(object = sce_LMC, group.by = "clusterLabel")

# Select database
CellChatDB <- CellChatDB.mouse
CellChatDB.use <- CellChatDB
cellchat_LMC@DB <- CellChatDB.use

# Run analysis
cellchat_LMC <- subsetData(cellchat_LMC)
cellchat_LMC <- identifyOverExpressedGenes(cellchat_LMC)
cellchat_LMC <- identifyOverExpressedInteractions(cellchat_LMC)

# Filter
options(future.globals.maxSize = 8000 * 1024^2) #necessary if limit too small
cellchat_LMC <- filterCommunication(cellchat_LMC, min.cells = 5)
cellchat_LMC <- computeCommunProb(cellchat_LMC, type =  "truncatedMean", trim = 0.2, population.size = FALSE)

cellchat_LMC <- computeCommunProbPathway(cellchat_LMC)

cellchat_LMC <- aggregateNet(cellchat_LMC)

future.seed=TRUE
cellchat_LMC <- netAnalysis_computeCentrality(cellchat_LMC, slot.name = "netP") 
```

# Save CellChat Ctrl
```{r, include = T, eval = F}
basedir <- here()
saveRDS(cellchat_LMC, file = paste0(basedir, "/data/pericarditis_allmerged_cleaned_cellchat_LMC.rds"))
```

# Run CellChat for TCRM
```{r, include = T, eval = F}
# Subset TCRM
seurat_TCRM <- subset(seuratM, cond %in% "TCRM")
sce_TCRM <- as.SingleCellExperiment(seurat_TCRM)
rownames(sce_TCRM) = gsub("^.*\\.","",rownames(sce_TCRM))

# Create CellChat object
cellchat_TCRM <- createCellChat(object = sce_TCRM, group.by = "clusterLabel")

# Select database
CellChatDB <- CellChatDB.mouse
CellChatDB.use <- CellChatDB
cellchat_TCRM@DB <- CellChatDB.use

# Run analysis
cellchat_TCRM <- subsetData(cellchat_TCRM)
cellchat_TCRM <- identifyOverExpressedGenes(cellchat_TCRM)
cellchat_TCRM <- identifyOverExpressedInteractions(cellchat_TCRM)

# Filter
options(future.globals.maxSize = 8000 * 1024^2) #necessary if limit too small
cellchat_TCRM <- filterCommunication(cellchat_TCRM, min.cells = 5)
cellchat_TCRM <- computeCommunProb(cellchat_TCRM, type =  "truncatedMean", trim = 0.2, population.size = FALSE)

cellchat_TCRM <- computeCommunProbPathway(cellchat_TCRM)

cellchat_TCRM <- aggregateNet(cellchat_TCRM)

future.seed=TRUE
cellchat_TCRM <- netAnalysis_computeCentrality(cellchat_TCRM, slot.name = "netP") 
```

# Save CellChat TCRM
```{r, include = T, eval = F}
basedir <- here()
saveRDS(cellchat_TCRM, file = paste0(basedir, "/data/pericarditis_allmerged_cleaned_cellchat_TCRM.rds"))
```



```{r, include = F, eval = T}
# Load CellChat files
basedir<- here()
fileNam <- paste0(basedir, "/data/pericarditis_allmerged_cleaned_cellchat_LMC.rds")
cellchat_LMC <- readRDS(fileNam)

basedir<- here()
fileNam <- paste0(basedir, "/data/pericarditis_allmerged_cleaned_cellchat_TCRM.rds")
cellchat_TCRM <- readRDS(fileNam)
```


# Interaction analysis of IL6 across conditions
## LMC
### Heatmap of sender/ receiver (selected pathways)
```{r}
#cellchat_all@netP$pathway

# Define pathways to analyze
pathways.show <- c("IL6")

# Compute centrality
cellchat_LMC <- netAnalysis_computeCentrality(cellchat_LMC, slot.name = "netP")

# Loop through each pathway and plot the network role
for (pathway in pathways.show) {
  netAnalysis_signalingRole_network(
    cellchat_LMC,
    signaling = pathway,
    width = 8,
    height = 2.5,
    font.size = 10,
    color.use = colPal,
    color.heatmap = "Greys"
  )
}
```

### Chord diagram for selected pathways
```{r}
# Define signaling pathways you to visualize
pathways.show <- c("IL6")


# Loop through each and generate the chord diagram
for (pathway in pathways.show) {
  netVisual_aggregate(
    cellchat_LMC,
    signaling = pathway,
    layout = "chord",
    color.use = colPal,
    title.name = pathway
  )
}
```


## TCRM
### Heatmap of sender/ receiver (selected pathways)
```{r}
#cellchat_all@netP$pathway

# Define pathways to analyze
pathways.show <- c("IL6")

# Compute centrality
cellchat_TCRM <- netAnalysis_computeCentrality(cellchat_TCRM, slot.name = "netP")

# Loop through each pathway and plot the network role
for (pathway in pathways.show) {
  netAnalysis_signalingRole_network(
    cellchat_TCRM,
    signaling = pathway,
    width = 8,
    height = 2.5,
    font.size = 10,
    color.use = colPal,
    color.heatmap = "Reds"
  )
}
```

### Chord diagram for selected pathways
```{r}
# Define signaling pathways you to visualize
pathways.show <- c("IL6")


# Loop through each and generate the chord diagram
for (pathway in pathways.show) {
  netVisual_aggregate(
    cellchat_TCRM,
    signaling = pathway,
    layout = "chord",
    color.use = colPal,
    title.name = pathway
  )
}
```

# Session info
```{r}
sessionInfo()
date()
```


