---
title: "Cellchat T cells and Fibroblasts"
author: "Nadine"
date: "2026-04-10"
output: 
html_document:
  self_contained: no
  code_folding: show
  number_sections: no
  toc: yes
  toc_depth: 3
  toc_float: true
  collapsed: no
  smooth_scroll: yes
editor_options:
  chunk_output_type: inline
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(warning = FALSE, message = FALSE, dev = c("png", "pdf"))
seed <- 1234
```


# Load packages
```{r}
suppressPackageStartupMessages({
  library(Seurat)
  library(ggpubr)
  library(dplyr)
  library(ggplot2)
  library(here)
  library(SingleCellExperiment)
  library(CellChat)
})
```


# Load seurat object
```{r}
# Load merged and cleaned object containing FBs
basedir<- here()
fileNam <- paste0(basedir, "/data/pericarditis_allmerged_FB_cleaned_cL_seurat.rds")
seuratFB <- readRDS(fileNam)
dim(seuratFB)

# Load merged and cleaned object containing T cells
basedir<- here()
fileNam <- paste0(basedir, "/data/pericarditis_allmerged_Tcells_cleaned_seurat_cL.rds")
seuratT <- readRDS(fileNam)
dim(seuratT)

```


# Transfer metadata
```{r}
seuratFB$celltype <- seuratFB$FBclusterLabel
seuratFB$lineage <- "Fibroblast"

seuratT$celltype <- seuratT$TclusterLabel
seuratT$lineage <- "Tcell"
```


# Merge seurat objects
```{r}
# Merge seurat objects with original identity
seuratM <- merge(
  x = seuratFB,
  y = seuratT,
  add.cell.ids = c("FB", "T"),
  project = "FB_T_Merged"
)

# Join layers
seuratM <- JoinLayers(seuratM)
```


# Metrics
```{r}
Idents(seuratM) <- seuratM$celltype

seuratM$fibroblast_label <- seuratM$FBclusterLabel
seuratM$tcell_label <- seuratM$TclusterLabel
table(seuratM$fibroblast_label)
table(seuratM$tcell_label)
table(seuratM$celltype)
table(seuratM$lineage)
```

# Filter
```{r}
# Remove CD4 Tem due to low cell count
seuratM <- subset(seuratM, celltype %in% c("CD4 Tcm"), invert=TRUE)
```


# Define color palette
```{r}
# Color palette celltype
colPal <- c( "#C1CDC1", "#EEE8AA", "#556B2F", "#CD6839", "#B2CD5B", "#A2B5CD","#DCA86C", "#6B7F99", "#CDCDC1",  "#8C8C82", "#A3B18A","#B5522A" ,  "#2F4F4F" , "#CD6600",  "#2B2B2B" ,"#8B3E2F"  , "#CDC673")

names(colPal) <- c("Fat4 Fibroblasts","Ccl11 Fibroblasts", "Tcf21 Fibroblasts", "Cxcl13 CoverCells", "Ccl19 FRCs", "Bmp4 Fibroblasts", "Col11 Fibroblasts", "Dpp4 Fibroblasts", "CD4 Tscm", "CD4 Tem", "Th1/Th17", "gdTcell-1", "CD8 Tscm", "gdTcell-2", "NKTcell", "CD8 Teff", "Tregs") 

```


# Set order of clusterLabel
```{r}
order <- c("Ccl19 FRCs", "Cxcl13 CoverCells", "Dpp4 Fibroblasts", "Ccl11 Fibroblasts", "Fat4 Fibroblasts", "Bmp4 Fibroblasts", "Tcf21 Fibroblasts", "Col11 Fibroblasts", "CD4 Tscm", "CD4 Tcm", "CD4 Tem", "CD4 Teff", "Tregs", "CD8 Tscm", "CD8 Teff", "gdTcell-1",  "gdTcell-2", "NKTcell") 

seuratM$celltype <- factor(seuratM$celltype, levels = order)
```



# Run CellChat for Ctrl
```{r, include = T, eval = F}
# Subset Controls
seurat_LMC <- subset(seuratM, subset = cond == "LMC")
sce_LMC <- as.SingleCellExperiment(seurat_LMC)

# Create CellChat object
cellchat_LMC <- createCellChat(object = sce_LMC, group.by = "celltype")

# Select database
CellChatDB <- CellChatDB.mouse
CellChatDB.use <- CellChatDB
cellchat_LMC@DB <- CellChatDB.use

# Run analysis
cellchat_LMC <- subsetData(cellchat_LMC)
cellchat_LMC <- identifyOverExpressedGenes(cellchat_LMC)
cellchat_LMC <- identifyOverExpressedInteractions(cellchat_LMC)

# Filter
options(future.globals.maxSize = 8000 * 1024^2)
cellchat_LMC <- filterCommunication(cellchat_LMC, min.cells = 5)
cellchat_LMC <- computeCommunProb(cellchat_LMC, type =  "truncatedMean", trim = 0.2, population.size = FALSE)

cellchat_LMC <- computeCommunProbPathway(cellchat_LMC)

cellchat_LMC <- aggregateNet(cellchat_LMC)

future.seed=TRUE
cellchat_LMC <- netAnalysis_computeCentrality(cellchat_LMC, slot.name = "netP") 
```

# Save CellChat Ctrl
```{r, include = T, eval = F}
basedir <- here()
saveRDS(cellchat_LMC, file = paste0(basedir, "/data/pericarditis_T_FB_merged_cleaned_cellchat_LMC.rds"))
```


# Run CellChat for TCRM
```{r, include = T, eval = F}
# Subset TCRM
seurat_TCRM <- subset(seuratM, subset = cond == "TCRM")
sce_TCRM <- as.SingleCellExperiment(seurat_TCRM)

# Create CellChat object
cellchat_TCRM <- createCellChat(object = sce_TCRM, group.by = "celltype")

# Select database
CellChatDB <- CellChatDB.mouse
CellChatDB.use <- CellChatDB
cellchat_TCRM@DB <- CellChatDB.use

# Run analysis
cellchat_TCRM <- subsetData(cellchat_TCRM)
cellchat_TCRM <- identifyOverExpressedGenes(cellchat_TCRM)
cellchat_TCRM <- identifyOverExpressedInteractions(cellchat_TCRM)

# Filter
options(future.globals.maxSize = 8000 * 1024^2)
cellchat_TCRM <- filterCommunication(cellchat_TCRM, min.cells = 5)
cellchat_TCRM <- computeCommunProb(cellchat_TCRM, type =  "truncatedMean", trim = 0.2, population.size = FALSE)

cellchat_TCRM <- computeCommunProbPathway(cellchat_TCRM)

cellchat_TCRM <- aggregateNet(cellchat_TCRM)

future.seed=TRUE
cellchat_TCRM <- netAnalysis_computeCentrality(cellchat_TCRM, slot.name = "netP") 
```

# Save CellChat TCRM
```{r, include = T, eval = F}
basedir <- here()
saveRDS(cellchat_TCRM, file = paste0(basedir, "/data/pericarditis_T_FB_merged_cleaned_cellchat_TCRM.rds"))
```


```{r, include = F, eval = T}
# Load CellChat files

## Ctrl
basedir<- here()
fileNam <- paste0(basedir, "/data/pericarditis_T_FB_merged_cleaned_cellchat_LMC.rds")
cellchat_LMC <- readRDS(fileNam)

## TCRM
basedir<- here()
fileNam <- paste0(basedir, "/data/pericarditis_T_FB_merged_cleaned_cellchat_TCRM.rds")
cellchat_TCRM <- readRDS(fileNam)
```


# Heatmap of IL-6 sender/ receiver {.tabset}

## Ctrl
```{r}
#cellchat_LMC@netP$pathway

# Define pathways to analyze
pathways.show <- c("IL6")

# Compute centrality
cellchat_LMC <- netAnalysis_computeCentrality(cellchat_LMC, slot.name = "netP")

# Loop through each pathway and plot the network role
for (pathway in pathways.show) {
  netAnalysis_signalingRole_network(
    cellchat_LMC,
    signaling = pathway,
    width = 8,
    height = 2.5,
    font.size = 10,
    color.use = colPal,
    color.heatmap = "Greys"
  )
}
```

## TCRM
```{r}
#cellchat_TCRM@netP$pathway

# Define pathways to analyze
pathways.show <- c("IL6")

# Compute centrality
cellchat_TCRM <- netAnalysis_computeCentrality(cellchat_TCRM, slot.name = "netP")

# Loop through each pathway and plot the network role
for (pathway in pathways.show) {
  netAnalysis_signalingRole_network(
    cellchat_TCRM,
    signaling = pathway,
    width = 8,
    height = 2.5,
    font.size = 10,
    color.use = colPal,
    color.heatmap = "Reds"
  )
}
```


# Chord diagram of IL-6 sender/ receiver {.tabset}
## Ctrl
```{r}
# Define signaling pathways you to visualize
pathways.show <- c("IL6")


# Loop through each and generate the chord diagram
for (pathway in pathways.show) {
  netVisual_aggregate(
    cellchat_LMC,
    signaling = pathway,
    layout = "chord",
    color.use = colPal,
    title.name = pathway
  )
}
```

## TCRM
```{r}
# Define signaling pathways you to visualize
pathways.show <- c("IL6")


# Loop through each and generate the chord diagram
for (pathway in pathways.show) {
  netVisual_aggregate(
    cellchat_TCRM,
    signaling = pathway,
    layout = "chord",
    color.use = colPal,
    title.name = pathway
  )
}
```


# Comparative analysis of phenotypes

# Merge cellchat objects
```{r, include = T, eval = F}
# Merge CellChat objects
object_list <- list(LMC = cellchat_LMC, TCRM = cellchat_TCRM)
cellchat_merged <- mergeCellChat(object_list, add.names = names(object_list))

cellchat_merged <- subsetData(cellchat_merged)

# Save merged CellChat object
basedir <- here()
saveRDS(cellchat_merged, file = paste0(basedir, "/data/pericarditis_T_FB_merged_cleaned_cellchat_cond_merged.rds"))
```



```{r, include=F, eval=T}
# Load merged cellchat file
basedir<- here()
fileNam <- paste0(basedir, "/data/pericarditis_T_FB_merged_cleaned_cellchat_cond_merged.rds")
cellchat_merged <- readRDS(fileNam)

object_list <- list(LMC = cellchat_LMC, TCRM = cellchat_TCRM)

```


# Bubble plots FRC - T cell signaling {.tabset}

## FRC to T cells by condition
```{r}
gg1 <- netVisual_bubble(
  cellchat_merged,
  sources.use = c("Ccl19 FRCs"),
  targets.use = c(
    "CD4 Tscm", "CD4 Tem", "CD4 Teff",
    "gdTcell-1", "CD8 Tscm", "gdTcell-2",
    "NKTcell", "CD8 Teff", "Tregs"
  ),
  signaling = c("CCL", "CXCL", "IL6", "ICAM"),
  comparison = c(1, 2),
  max.dataset = 2,
  title.name = "Increased Chemokine + IL6 signaling in TCRM",
  angle.x = 45,
  remove.isolate = TRUE,
  color.text = c("lightgray", "#B45B5C")
)
gg1

```

## FRC to ab T cells by condition
```{r}
gg1 <- netVisual_bubble(
  cellchat_merged,
  sources.use = c("Ccl19 FRCs"),
  targets.use = c(
    "CD4 Tscm", "CD4 Tem", "CD4 Teff",
   "CD8 Tscm", "CD8 Teff", "Tregs"
  ),
  signaling = c("CCL", "CXCL", "IL6", "ICAM"),
  comparison = c(1, 2),
  max.dataset = 2,
  title.name = "Increased Chemokine + IL6 signaling in TCRM",
  angle.x = 45,
  remove.isolate = TRUE,
  color.text = c("lightgray", "#B45B5C")
)
gg1

```


# Session info
```{r}
sessionInfo()
date()
```

