---
title: "characterize full dataset"
author: "Nadine"
date: "2025-04-09"
output: 
html_document:
  self_contained: no
  code_folding: show
  number_sections: no
  toc: yes
  toc_depth: 3
  toc_float: true
  collapsed: no
  smooth_scroll: yes
editor_options:
  chunk_output_type: inline
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(warning = FALSE, message = FALSE, dev = c("png", "pdf"))
seed <- 1234
```

#Load packages
```{r}
suppressPackageStartupMessages({
  library(dplyr)
  library(reshape2)
  library(ggplot2)
  library(cowplot)
  library(purrr)
  library(Seurat)
  library(tidyverse)
  library(ggpubr)
  library(runSeurat3)
  library(here)
  library(ggsci)
  library(pheatmap)
  library(scater)
})
```


#Load seurat object
```{r}
#Load merged and cleaned object from previous analysis
basedir<- here()
fileNam <- paste0(basedir, "/data/pericarditis_allmerged_cleaned_seurat.rds")
seuratM <- readRDS(fileNam)
```

#Marker genes per cluster
```{r, include = T, eval = T}
#Set identity to resolution 0.25
Idents(seuratM) <-  seuratM$RNA_snn_res.0.25

#Run Marker Genes
seuratM_markers_all <- FindAllMarkers(object = seuratM, assay ="RNA",
                                     only.pos = TRUE, min.pct = 0.25,
                                     logfc.threshold = 0.25,
                                     test.use = "wilcox")

#Save Marker Genes as Text File
basedir<- here()
basedir <- paste0(basedir, "/docs/")
write.table(
  seuratM_markers_all,
  file = paste0(basedir, "pericarditis_allmerged_0.25_cleaned_markerGenes.txt"),
  row.names = FALSE,
  quote = FALSE,
  sep = "\t"
)

```

#Assign labels to clusters
```{r}
seuratM$cluster <- seuratM$RNA_snn_res.0.25

seuratM$clusterLabel <- "clusterLabel"
seuratM$clusterLabel[which(seuratM$RNA_snn_res.0.25 %in% c("0"))] <- "B2 Bcells"
seuratM$clusterLabel[which(seuratM$RNA_snn_res.0.25 %in% c("3"))] <- "B1 Bcells"
seuratM$clusterLabel[which(seuratM$RNA_snn_res.0.25 %in% c("10"))] <- "Plasma cells"

seuratM$clusterLabel[which(seuratM$RNA_snn_res.0.25 %in% c("1"))] <- "Fibroblasts-1"
seuratM$clusterLabel[which(seuratM$RNA_snn_res.0.25 %in% c("6"))] <- "Cover cells"
seuratM$clusterLabel[which(seuratM$RNA_snn_res.0.25 %in% c("4"))] <- "Mesothelial cells"
seuratM$clusterLabel[which(seuratM$RNA_snn_res.0.25 %in% c("13"))] <- "Fibroblasts-2"

seuratM$clusterLabel[which(seuratM$RNA_snn_res.0.25 %in% c("2"))] <- "Tcells"
seuratM$clusterLabel[which(seuratM$RNA_snn_res.0.25 %in% c("8"))] <- "gdTcells"
seuratM$clusterLabel[which(seuratM$RNA_snn_res.0.25 %in% c("12"))] <- "ILCs"


seuratM$clusterLabel[which(seuratM$RNA_snn_res.0.25 %in% c("7"))] <- "Mono/Mphs"
seuratM$clusterLabel[which(seuratM$RNA_snn_res.0.25 %in% c("9"))] <- "Neutrophils"
seuratM$clusterLabel[which(seuratM$RNA_snn_res.0.25 %in% c("14"))] <- "DCs"
seuratM$clusterLabel[which(seuratM$RNA_snn_res.0.25 %in% c("5"))] <- "NKcells"


seuratM$clusterLabel[which(seuratM$RNA_snn_res.0.25 %in% c("11"))] <- "Cycling cells"
seuratM$clusterLabel[which(seuratM$RNA_snn_res.0.25 %in% c("15"))] <- "Adipocytes"
```


#Set color palette for clusters
```{r}
colPal <- c( "#CDC1C5","#8B7B8B","#CDC673","#CD5B45","#C1CDC1","#7AC5CD","#7F7F7F","#4A708B","#CD919E","#8B668B","#8B475D","#8FBC8F","#CCCCCC","#8B7D6B","#CDC8B1","#333333")

names(colPal) <- c("B2 Bcells","B1 Bcells", "Cycling cells","Plasma cells", "Tcells", "ILCs", "gdTcells",  "NKcells", "Mono/Mphs", "Neutrophils", "DCs", "Fibroblasts-1", "Fibroblasts-2", "Cover cells","Mesothelial cells", "Adipocytes") 

colPal_Phenotype <- c("#8B0000", "#FCDACB", "#E6E6E6")
names(colPal_Phenotype) <- c( "iCMP", "inflamed", "healthy")

```

#Stats
```{r}
table(seuratM$orig.ident)
table(seuratM$clusterLabel)
table(seuratM$clusterLabel, seuratM$cond)
table(seuratM$Phenotype)
table(seuratM$clusterLabel, seuratM$Phenotype)
table(seuratM$cond)
```

#Visualization in umaps
```{r}
Idents(seuratM) <-  seuratM$clusterLabel
DimPlot(seuratM, reduction = "umap", cols=colPal, group.by= "clusterLabel")+
  theme_void()

DimPlot(seuratM, reduction = "umap", cols=colPal, label = T, group.by= "clusterLabel")+
  theme_void()


DimPlot(seuratM, reduction = "umap", cols=c("#D3D3D3", "#B45B5C"), group.by = "cond")+
  theme_void()


DimPlot(seuratM, reduction = "umap", cols=c("#D3D3D3", "#B45B5C"), group.by = "cond", split.by = "cond")+
  theme_void()

DimPlot(seuratM, reduction = "umap", cols=c("#99B898", "#628395"), group.by = "age", split.by = "age")+
  theme_void()

DimPlot(seuratM, reduction = "umap", cols=colPal_Phenotype, group.by = "Phenotype")+
  theme_void()
```

#Calculate marker genes
```{r, include = T, eval = T}

#Set identity to set clusterLabels
Idents(seuratM) <-  seuratM$clusterLabel

#Run Marker Genes
seuratM_markers_all <- FindAllMarkers(object = seuratM, assay ="RNA",
                                     only.pos = TRUE, min.pct = 0.25,
                                     logfc.threshold = 0.25,
                                     test.use = "wilcox")

#Save Marker Genes as Text File
basedir<- here()
basedir <- paste0(basedir, "/docs/")
write.table(
  seuratM_markers_all,
  file = paste0(basedir, "pericarditis_allmerged_0.25_cL_cleaned_markerGenes.txt"),
  row.names = FALSE,
  quote = FALSE,
  sep = "\t"
)


```

#Load marker genes
```{r}

basedir<- here()
fileNam <- paste0(basedir, "/docs/pericarditis_allmerged_0.25_cL_cleaned_markerGenes.txt")
seuratM_MarkerGenes <- read.table(fileNam, header = TRUE, sep = "\t", stringsAsFactors = FALSE)
  
```

#Load average heatmap function
```{r avg heatmap funct}

avgHeatmap <- function(seuratA, selGenes, colVecIdent, colVecCond=NULL,
                       ordVec=NULL, gapVecR=NULL, gapVecC=NULL,cc=FALSE,
                       cr=FALSE, condCol=FALSE){
  
  selGenes <- selGenes$gene
  
  ## assay data
  clusterAssigned <- as.data.frame(Idents(seuratA)) %>%
  dplyr::mutate(cell=rownames(.))
  colnames(clusterAssigned)[1] <- "ident"
  seuratADat <- GetAssayData(seuratA)
  
  ## genes of interest
  genes <- data.frame(gene=rownames(seuratA)) %>% 
    mutate(geneID=gsub("^.*\\.", "", gene)) %>% filter(geneID %in% selGenes)

  ## matrix with averaged cnts per ident
  logNormExpres <- as.data.frame(t(as.matrix(
    seuratADat[which(rownames(seuratADat) %in% genes$gene),])))
  logNormExpres <- logNormExpres %>% dplyr::mutate(cell=rownames(.)) %>%
    dplyr::left_join(.,clusterAssigned, by=c("cell")) %>%
    dplyr::select(-cell) %>% dplyr::group_by(ident) %>%
    dplyr::summarise_all(mean)
  logNormExpresMa <- logNormExpres %>% dplyr::select(-ident) %>% as.matrix()
  rownames(logNormExpresMa) <- logNormExpres$ident
  logNormExpresMa <- t(logNormExpresMa)
  rownames(logNormExpresMa) <- gsub("^.*?\\.","",rownames(logNormExpresMa))
  
  ## remove genes if they are all the same in all groups
  ind <- apply(logNormExpresMa, 1, sd) == 0
  logNormExpresMa <- logNormExpresMa[!ind,]
  genes <- genes[!ind,]

  ## color columns according to cluster
  annotation_col <- as.data.frame(gsub("(^.*?_)","",
                                       colnames(logNormExpresMa)))%>%
    dplyr::mutate(celltype=gsub("(_.*$)","",colnames(logNormExpresMa)))
  colnames(annotation_col)[1] <- "col1"
  annotation_col <- annotation_col %>%
    dplyr::mutate(cond = gsub("(^[0-9]_?)","",col1)) %>%
    dplyr::select(cond, celltype)
  rownames(annotation_col) <- colnames(logNormExpresMa) 

  ann_colors = list(
      #cond = colVecCond,
      celltype=colVecIdent)
  if(is.null(ann_colors$cond)){
    annotation_col$cond <- NULL
  }
  
  ## adjust order
  logNormExpresMa <- logNormExpresMa[selGenes,]
  if(is.null(ordVec)){
    ordVec <- levels(seuratA)
  }
  logNormExpresMa <- logNormExpresMa[,ordVec]

  ## scaled row-wise
  pheatmap(logNormExpresMa, scale="row" ,treeheight_row = 0, cluster_rows = cr, 
         cluster_cols = cc,
         color = colorRampPalette(c("#2166AC", "#F7F7F7", "#B2182B"))(50),
         annotation_col = annotation_col, cellwidth=15, cellheight=10,
         annotation_colors = ann_colors, gaps_row = gapVecR, gaps_col = gapVecC)
}
```

#Heatmap of top 10 marker genes per cluster
```{r, fig.height=21}
cluster <- levels(seuratM)
selGenesAll <- seuratM_MarkerGenes %>% group_by(cluster) %>% 
  top_n(-8, p_val_adj) %>% 
  top_n(8, avg_log2FC)
names(selGenesAll) [names(selGenesAll) == "gene"] <- "Gene"
selGenesAll <- selGenesAll %>% mutate(gene=gsub("^.*\\.", "", Gene)) %>% filter(nchar(Gene)>1)

#Set color palette
colPal <- c( "#CDC1C5","#8B7B8B","#CDC673","#CD5B45","#C1CDC1","#7AC5CD","#7F7F7F","#4A708B","#CD919E","#8B668B","#8B475D","#8FBC8F","#CCCCCC","#8B7D6B","#CDC8B1","#333333")
names(colPal) <- c("B2 Bcells","B1 Bcells", "Cycling cells","Plasma cells", "Tcells", "ILCs", "gdTcells",  "NKcells", "Mono/Mphs", "Neutrophils", "DCs", "Fibroblasts-1", "Fibroblasts-2", "Cover cells","Mesothelial cells", "Adipocytes")

#Set order
ordVec <- names(colPal) 


Idents(seuratM) <- seuratM$clusterLabel
pOut <- avgHeatmap(seurat = seuratM, selGenes = selGenesAll,
                  colVecIdent = colPal, 
                  ordVec=ordVec,
                  gapVecR=NULL, gapVecC=NULL,cc=FALSE,
                  cr=T, condCol=F)
```

#Heatmap of selected marker genes
```{r, fig.height=7}
selMarker <- list(Bcells_ = c("Ptprc", "Cd19", "Pax5", "Ighm", "Ighd"),                
                  Tcells_ = c("Cd3d", "Cd3e", "Cd4", "Cd8a", "Cd5"),
                  Innate_ = c("Il23r", "Zbtb16", "Nkg7", "Eomes", "Gata3", "Il7r"),                  
                  Myeloids_ = c("Itgam", "Cd14", "Ly6g","Siglech", "Itgax", "Cd68"),
                  Stroma_ = c("Pdgfra", "Col1a1", "Pdpn","Cxcl13", "Cd34", "Mcam", "Msln", "Clu", "Krt19", "Adipoq"))
    
Idents(seuratM) <- seuratM$clusterLabel

selGenes <- data.frame(gene=unlist(selMarker)) %>%
  rownames_to_column(var="grp") %>% mutate(Grp=gsub("_.*", "", grp))
grpCnt <- selGenes %>% group_by(Grp) %>% summarise(cnt=n())
gapR <- data.frame(Grp=unique(selGenes$Grp)) %>% 
  left_join(.,grpCnt, by="Grp") %>% mutate(cumSum=cumsum(cnt)) 

pOut <- avgHeatmap(seurat = seuratM, selGenes = selGenes,
                  colVecIdent = colPal, 
                  ordVec=ordVec,
                  gapVecR=gapR$cumSum, gapVecC=NULL,cc=T,
                  cr=T, condCol=F)


```

#Abundance across all samples
```{r}
#table(seuratM$SampleName)

#Plot abundance of clusters in individual samples
datList <- NULL
for(con in unique(seuratM$SampleName)){
  seuratMSub <- subset(seuratM, SampleName==con)
  print(dim(seuratMSub))
  dat_SampleName <- as.data.frame(table(seuratMSub$clusterLabel)) %>%
  mutate(percent=Freq/ncol(seuratMSub)) %>% mutate(SampleName=con)
  datList[[con]] <- dat_SampleName
}
dat_all <- do.call("rbind", datList)


ordX <- c("4w_LMC1", "4w_LMC2", "4w_LMC3", "4w_LMC4", "8w_LMC1", "8w_LMC2", "8w_LMC3", "8w_LMC4", "8w_LMC5", "8w_LMC6", "8w_LMC7", "8w_LMC8" , "4w_TCRM1", "4w_TCRM2", "4w_TCRM3", "4w_TCRM4", "4w_TCRM5", "4w_TCRM6", "4w_TCRM7", "8w_TCRM1", "8w_TCRM2", "8w_TCRM3", "8w_TCRM4", "8w_TCRM5", "8w_TCRM6", "8w_TCRM7", "8w_TCRM8")

#Create barplot with samples on x axis and cluster colors
ggbarplot(dat_all, x= "SampleName", y= "percent", fill = "Var1", legend = "right", legend.titel = "cluster", ylab = "frequency") +  scale_fill_manual(values = colPal) + theme(axis.text.x = element_text(angle = 90)) + scale_x_discrete(limits=ordX) 



```


#Abundance across non-sorted samples
```{r}
#Plot abundance of clusters in individual samples
##Leave out sorted samples

seuratMSub <- subset(seuratM, SampleName %in% c("8w_LMC8", "8w_TCRM7", "8w_TCRM8"), invert=TRUE )
  
##run calculations
datList <- NULL
for(con in unique(seuratMSub$SampleName)){
  seuratM_sub <- subset(seuratMSub, SampleName==con)
  print(dim(seuratM_sub))
  dat_SampleName <- as.data.frame(table(seuratM_sub$clusterLabel)) %>%
  mutate(percent=Freq/ncol(seuratM_sub)) %>% mutate(SampleName=con)
  datList[[con]] <- dat_SampleName
}
dat_all <- do.call("rbind", datList)


ordX <- c("4w_LMC1", "4w_LMC2", "4w_LMC3", "4w_LMC4", "8w_LMC1", "8w_LMC2", "8w_LMC3", "8w_LMC4", "8w_LMC5", "8w_LMC6", "8w_LMC7",  "4w_TCRM1", "4w_TCRM2", "4w_TCRM3", "4w_TCRM4", "4w_TCRM5", "4w_TCRM6", "4w_TCRM7", "8w_TCRM1", "8w_TCRM2", "8w_TCRM3", "8w_TCRM4", "8w_TCRM5", "8w_TCRM6")

#Create barplot with samples on x axis and cluster colors
ggbarplot(dat_all, x= "SampleName", y= "percent", fill = "Var1", legend = "right", legend.titel = "cluster", ylab = "frequency") +  scale_fill_manual(values = colPal) + theme(axis.text.x = element_text(angle = 90)) + scale_x_discrete(limits=ordX) 



```

#Abundance across phenotypes
```{r}

datList <- NULL
for(con in unique(seuratM$Phenotype)){
  seuratMSub <- subset(seuratM, Phenotype==con)
  print(dim(seuratMSub))
  dat_Phenotype <- as.data.frame(table(seuratMSub$clusterLabel)) %>%
  mutate(percent=Freq/ncol(seuratMSub)) %>% mutate(Phenotype=con)
  datList[[con]] <- dat_Phenotype
}
dat_all <- do.call("rbind", datList)

#Set order of Phenotypes
ordX <- c("healthy", "inflamed", "iCMP")

#Create barplot with Phenotypes on x axis and cluster colors
ggbarplot(dat_all, x= "Phenotype", y= "percent", fill = "Var1", legend = "right", legend.titel = "cluster", ylab = "frequency") +  scale_fill_manual(values = colPal) + theme(axis.text.x = element_text(angle = 90)) + scale_x_discrete(limits=ordX) 
```



#Save seurat object
```{r}
#Save seurat object
basedir <- here()
saveRDS(seuratM, file = paste0(basedir, "/data/pericarditis_allmerged_cleaned_cL_seurat.rds"))
```


#Session info
```{r}
sessionInfo()
date()
```

