---

title: "Fibroblasts TCRM and Ctrl"
author: "Nadine"
date: "2026-04-15"
output: 
html_document:
  self_contained: no
  code_folding: show
  number_sections: no
  toc: yes
  toc_depth: 3
  toc_float: true
  collapsed: no
  smooth_scroll: yes
editor_options:
  chunk_output_type: inline
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(warning = FALSE, message = FALSE, dev = c("png", "pdf"))
seed <- 1234
```

#Load packages
```{r}
suppressPackageStartupMessages({
  library(Seurat)
  library(ggpubr)
  library(pheatmap)
  library(dplyr)
  library(tidyverse)
  library(clusterProfiler)
  library(org.Mm.eg.db)
  library(enrichplot)
  library(DOSE)
  library(grid)
  library(ggplot2)
  library(here)
  library (tibble)
})
```


# Load seurat object
```{r}
# Load merged and cleaned object containing FBs
basedir<- here()
fileNam <- paste0(basedir, "/data/pericarditis_allmerged_FB_cleaned_seurat.rds")
seuratFB <- readRDS(fileNam)
```


## DimPlot Clusters {.tabset}

### Resolution 0.2
```{r}
Idents(seuratFB) <-  seuratFB$RNA_snn_res.0.2
DimPlot(seuratFB, reduction = "umap", label = T)
```

### Resolution 0.25
```{r}
Idents(seuratFB) <-  seuratFB$RNA_snn_res.0.25
DimPlot(seuratFB, reduction = "umap", label = T)
```

### Resolution 0.4
```{r}
Idents(seuratFB) <-  seuratFB$RNA_snn_res.0.4
DimPlot(seuratFB, reduction = "umap", label = T)
```


# Marker genes per cluster
```{r, include = T, eval = F}
# Set identity to resolution 0.25
Idents(seuratFB) <-  seuratFB$RNA_snn_res.0.25

# Run marker genes
seuratFB_markers_all <- FindAllMarkers(object = seuratFB, assay ="RNA",
                                     only.pos = TRUE, min.pct = 0.25,
                                     logfc.threshold = 0.25,
                                     test.use = "wilcox")

# Save marker genes as text file
basedir<- here()
basedir <- paste0(basedir, "/docs/")
write.table(
  seuratFB_markers_all,
  file = paste0(basedir, "pericarditis_allmerged_FB_cleaned_0.25_markerGenes.txt"),
  row.names = FALSE,
  quote = FALSE,
  sep = "\t"
)
```


# Assign labels to fibroblast clusters
```{r}
seuratFB$cluster <- seuratFB$RNA_snn_res.0.25

seuratFB$FBclusterLabel <- "FBclusterLabel"
seuratFB$FBclusterLabel[which(seuratFB$RNA_snn_res.0.25 %in% c("0"))] <- "Fat4 Fibroblasts"
seuratFB$FBclusterLabel[which(seuratFB$RNA_snn_res.0.25 %in% c("1"))] <- "Ccl19 FRCs"
seuratFB$FBclusterLabel[which(seuratFB$RNA_snn_res.0.25 %in% c("2"))] <- "Tcf21 Fibroblasts"
seuratFB$FBclusterLabel[which(seuratFB$RNA_snn_res.0.25 %in% c("3"))] <- "Cxcl13 CoverCells"
seuratFB$FBclusterLabel[which(seuratFB$RNA_snn_res.0.25 %in% c("4"))] <- "Ccl11 Fibroblasts"
seuratFB$FBclusterLabel[which(seuratFB$RNA_snn_res.0.25 %in% c("5"))] <- "Dpp4 Fibroblasts"
seuratFB$FBclusterLabel[which(seuratFB$RNA_snn_res.0.25 %in% c("6"))] <- "Bmp4 Fibroblasts"
seuratFB$FBclusterLabel[which(seuratFB$RNA_snn_res.0.25 %in% c("7"))] <- "Col11 Fibroblasts"
```


# Define color palette
```{r}
# Color palette FBclusterLabel
colPal_FB <- c("#C1CDC1", "#EEE8AA", "#556B2F", "#CD6839", "#B2CD5B", "#A2B5CD","#DCA86C", "#6B7F99")
names(colPal_FB) <- c("Fat4 Fibroblasts","Ccl11 Fibroblasts", "Tcf21 Fibroblasts", "Cxcl13 CoverCells", "Ccl19 FRCs", "Bmp4 Fibroblasts", "Col11 Fibroblasts", "Dpp4 Fibroblasts")

# Color palette Phenotypes
colPal_Phenotype <- c("#8B0000", "#FCDACB", "#E6E6E6")
names(colPal_Phenotype) <- c( "iCMP", "inflamed", "healthy")
```


# Metrics
```{r}
table(seuratFB$orig.ident)
table(seuratFB$cond)
table(seuratFB$Phenotype)
table(seuratFB$FBclusterLabel)
```


## DimPlot Clusters {.tabset}

### FBclusterLabel
```{r}
Idents(seuratFB) <-  seuratFB$FBclusterLabel
DimPlot(seuratFB, reduction = "umap", cols=colPal_FB, group.by= "FBclusterLabel")
```

### Condition split
```{r}
DimPlot(seuratFB, reduction = "umap", cols=colPal_FB, label = F, group.by= "FBclusterLabel", split.by = "cond")
```



# Calculate marker genes per cluster
```{r, include = T, eval = F}
# Set identity to resolution FBclusterLabel
Idents(seuratFB) <-  seuratFB$FBclusterLabel

# Run marker genes
seuratFB_markers_all <- FindAllMarkers(object = seuratFB, assay ="RNA",
                                     only.pos = TRUE, min.pct = 0.25,
                                     logfc.threshold = 0.25,
                                     test.use = "wilcox")

# Save marker genes as text file
basedir<- here()
basedir <- paste0(basedir, "/docs/")
write.table(
  seuratFB_markers_all,
  file = paste0(basedir, "pericarditis_allmerged_FB_0.25_cL_markerGenes.txt"),
  row.names = FALSE,
  quote = FALSE,
  sep = "\t"
)
```


# Load marker genes text file
```{r, include=T, eval=T}
basedir<- here()
fileNam <- paste0(basedir, "/docs/pericarditis_allmerged_FB_0.25_cL_markerGenes.txt")
seuratFB_MarkerGenes <- read.table(fileNam, header = TRUE, sep = "\t", stringsAsFactors = FALSE)
```


# Load average heatmap function
```{r avg heatmap funct}

avgHeatmap <- function(seuratFB, selGenes, colVecIdent, colVecCond=NULL,
                       ordVec=NULL, gapVecR=NULL, gapVecC=NULL,cc=FALSE,
                       cr=FALSE, condCol=FALSE){
  
  selGenes <- selGenes$gene
  
  ## assay data
  clusterAssigned <- as.data.frame(Idents(seuratFB)) %>%
  dplyr::mutate(cell=rownames(.))
  colnames(clusterAssigned)[1] <- "ident"
  seuratFBDat <- GetAssayData(seuratFB)
  
  ## genes of interest
  genes <- data.frame(gene=rownames(seuratFB)) %>% 
    mutate(geneID=gsub("^.*\\.", "", gene)) %>% filter(geneID %in% selGenes)

  ## matrix with averaged cnts per ident
  logNormExpres <- as.data.frame(t(as.matrix(
    seuratFBDat[which(rownames(seuratFBDat) %in% genes$gene),])))
  logNormExpres <- logNormExpres %>% dplyr::mutate(cell=rownames(.)) %>%
    dplyr::left_join(.,clusterAssigned, by=c("cell")) %>%
    dplyr::select(-cell) %>% dplyr::group_by(ident) %>%
    dplyr::summarise_all(mean)
  logNormExpresMa <- logNormExpres %>% dplyr::select(-ident) %>% as.matrix()
  rownames(logNormExpresMa) <- logNormExpres$ident
  logNormExpresMa <- t(logNormExpresMa)
  rownames(logNormExpresMa) <- gsub("^.*?\\.","",rownames(logNormExpresMa))
  
  ## remove genes if they are all the same in all groups
  ind <- apply(logNormExpresMa, 1, sd) == 0
  logNormExpresMa <- logNormExpresMa[!ind,]
  genes <- genes[!ind,]

  ## color columns according to cluster
  annotation_col <- as.data.frame(gsub("(^.*?_)","",
                                       colnames(logNormExpresMa)))%>%
    dplyr::mutate(celltype=gsub("(_.*$)","",colnames(logNormExpresMa)))
  colnames(annotation_col)[1] <- "col1"
  annotation_col <- annotation_col %>%
    dplyr::mutate(cond = gsub("(^[0-9]_?)","",col1)) %>%
    dplyr::select(cond, celltype)
  rownames(annotation_col) <- colnames(logNormExpresMa) 

  ann_colors = list(
      #cond = colVecCond,
      celltype=colVecIdent)
  if(is.null(ann_colors$cond)){
    annotation_col$cond <- NULL
  }
  
  ## adjust order
  logNormExpresMa <- logNormExpresMa[selGenes,]
  if(is.null(ordVec)){
    ordVec <- levels(seuratFB)
  }
  logNormExpresMa <- logNormExpresMa[,ordVec]

  ## scaled row-wise
  pheatmap(logNormExpresMa, scale="row" ,treeheight_row = 0, cluster_rows = cr, 
         cluster_cols = cc,
         color = colorRampPalette(c("#2166AC", "#F7F7F7", "#B2182B"))(50),
         annotation_col = annotation_col, cellwidth=15, cellheight=10,
         annotation_colors = ann_colors, gaps_row = gapVecR, gaps_col = gapVecC)
}
```


## Heatmaps {.tabset}

### Top 10 marker genes per cluster
```{r, fig.height=15}
cluster <- levels(seuratFB)
selGenesAll <- seuratFB_MarkerGenes %>% group_by(cluster) %>% 
  top_n(-10, p_val_adj) %>% 
  top_n(10, avg_log2FC)
names(selGenesAll) [names(selGenesAll) == "gene"] <- "Gene"
selGenesAll <- selGenesAll %>% mutate(gene=gsub("^.*\\.", "", Gene)) %>% filter(nchar(Gene)>1)

colPal_FB <- c("#C1CDC1", "#EEE8AA", "#556B2F", "#CD6839", "#B2CD5B", "#A2B5CD","#DCA86C", "#6B7F99")
names(colPal_FB) <- c("Fat4 Fibroblasts","Ccl11 Fibroblasts", "Tcf21 Fibroblasts", "Cxcl13 CoverCells", "Ccl19 FRCs", "Bmp4 Fibroblasts", "Col11 Fibroblasts", "Dpp4 Fibroblasts")

ordVec <- c("Ccl19 FRCs", "Cxcl13 CoverCells", "Dpp4 Fibroblasts", "Ccl11 Fibroblasts", "Fat4 Fibroblasts", "Bmp4 Fibroblasts", "Tcf21 Fibroblasts", "Col11 Fibroblasts")

Idents(seuratFB) <- seuratFB$FBclusterLabel
pOut <- avgHeatmap(seurat = seuratFB, selGenes = selGenesAll,
                  colVecIdent = colPal_FB, 
                  ordVec=ordVec,
                  gapVecR=NULL, gapVecC=NULL,cc=FALSE,
                  cr=T, condCol=F)

```

### Selected immune interaction genes
```{r, fig.height=4, fig.width=5}

genes <- data.frame(gene=rownames(seuratFB))

selMarker <- list(stroma = c("Ccl19","Il6","Ccl2", "Cxcl1","Cxcl13", "Icam1", "Vcam1", "Il33", "Ccl11"))
    
Idents(seuratFB) <- seuratFB$FBclusterLabel

selGenes <- data.frame(gene=unlist(selMarker)) %>%
  rownames_to_column(var="grp") %>% mutate(Grp=gsub("_.*", "", grp))
grpCnt <- selGenes %>% group_by(Grp) %>% summarise(cnt=n())
gapR <- data.frame(Grp=unique(selGenes$Grp)) %>% 
  left_join(.,grpCnt, by="Grp") %>% mutate(cumSum=cumsum(cnt)) 

colPal_FB <- c("#C1CDC1", "#EEE8AA", "#556B2F", "#CD6839", "#B2CD5B", "#A2B5CD","#DCA86C", "#6B7F99")
names(colPal_FB) <- c("Fat4 Fibroblasts","Ccl11 Fibroblasts", "Tcf21 Fibroblasts", "Cxcl13 CoverCells", "Ccl19 FRCs", "Bmp4 Fibroblasts", "Col11 Fibroblasts", "Dpp4 Fibroblasts")

ordVec <- c("Ccl19 FRCs", "Cxcl13 CoverCells", "Dpp4 Fibroblasts", "Ccl11 Fibroblasts", "Fat4 Fibroblasts", "Bmp4 Fibroblasts", "Tcf21 Fibroblasts", "Col11 Fibroblasts")

pOut <- avgHeatmap(seuratFB = seuratFB, selGenes = selGenes,
                  colVecIdent = colPal_FB, 
                  ordVec=ordVec,
                  gapVecR=NULL, gapVecC=NULL,cc=F,
                  cr=T, condCol=F)
```

### Selected ECM genes
```{r, fig.height=4, fig.width=5}

genes <- data.frame(gene=rownames(seuratFB))

selMarker <- list(stroma = c("Col1a1", "Col15a1", "Col11a1", "Eln", "Col8a1", "Col6a1", "Tgfbi", "Dcn", "Ogn"))
    
Idents(seuratFB) <- seuratFB$FBclusterLabel

selGenes <- data.frame(gene=unlist(selMarker)) %>%
  rownames_to_column(var="grp") %>% mutate(Grp=gsub("_.*", "", grp))
grpCnt <- selGenes %>% group_by(Grp) %>% summarise(cnt=n())
gapR <- data.frame(Grp=unique(selGenes$Grp)) %>% 
  left_join(.,grpCnt, by="Grp") %>% mutate(cumSum=cumsum(cnt)) 

colPal_FB <- c("#C1CDC1", "#EEE8AA", "#556B2F", "#CD6839", "#B2CD5B", "#A2B5CD","#DCA86C", "#6B7F99")
names(colPal_FB) <- c("Fat4 Fibroblasts","Ccl11 Fibroblasts", "Tcf21 Fibroblasts", "Cxcl13 CoverCells", "Ccl19 FRCs", "Bmp4 Fibroblasts", "Col11 Fibroblasts", "Dpp4 Fibroblasts")

ordVec <- c("Ccl19 FRCs", "Cxcl13 CoverCells", "Dpp4 Fibroblasts", "Ccl11 Fibroblasts", "Fat4 Fibroblasts", "Bmp4 Fibroblasts", "Tcf21 Fibroblasts", "Col11 Fibroblasts")

pOut <- avgHeatmap(seuratFB = seuratFB, selGenes = selGenes,
                  colVecIdent = colPal_FB, 
                  ordVec=ordVec,
                  gapVecR=NULL, gapVecC=NULL,cc=F,
                  cr=T, condCol=F)
```
## Dotplot of selected marker genes {.tabset}

### Dotplot of selected marker genes (unordered)
```{r, fig.height=4, fig.width=5}

Idents(seuratFB) <- seuratFB$FBclusterLabel

# Marker genes as simple vector
selMarker <-  c("Pdgfra", "Pdpn", "Cxcl13", "Ccl19", "Ccl11",  "Tcf21",  "Bmp4", "Col11a1", "Dpp4", "Fat4", "Aldh1a2")

# Assign a full gene name to the short geneID
genes <- data.frame(gene=rownames(seuratFB)) %>% mutate(geneID=gsub(".*\\.", "", gene)) %>% 
  filter(geneID %in% as.character(selMarker)) %>% arrange(factor(geneID, levels = rev(selMarker)))

# Create dataframe of cL
clusterDat <- data.frame(cluster=levels(seuratFB)) %>%
  mutate(label = levels(seuratFB))


# Scale the gene across all clusters
DotPlot(
  object=seuratFB,
  features = genes$gene,
  cols = c("#2C3E50", "#F1E6A1"),
  col.min = -2.5,
  col.max = 2.5,
  dot.min = 0,
  dot.scale = 6,
  idents = NULL,
  group.by = NULL,
  split.by = NULL,
  cluster.idents = T, #TRUE= automatic order, FALSE= ordered 
  scale = T,
  scale.by = "radius",
  scale.min = NA,
  scale.max = NA
) + coord_flip() + scale_x_discrete(breaks=genes$gene, labels=genes$geneID)  + scale_y_discrete(breaks=clusterDat$cluster, labels=clusterDat$label) + theme(axis.text.x = element_text(angle = 90))

```

### Dotplot of selected marker genes (ordered)
```{r, fig.height=4, fig.width=5}

# Set cluster order
ordVec <- c("Ccl19 FRCs", "Cxcl13 CoverCells", "Dpp4 Fibroblasts", "Ccl11 Fibroblasts", "Fat4 Fibroblasts", "Bmp4 Fibroblasts", "Tcf21 Fibroblasts", "Col11 Fibroblasts")

# Set identities
Idents(seuratFB) <- seuratFB$FBclusterLabel
Idents(seuratFB) <- factor(Idents(seuratFB), levels = ordVec)

# Marker genes as simple vector
selMarker <-  c("Pdgfra", "Pdpn", "Cxcl13", "Ccl19", "Ccl11",  "Tcf21",  "Bmp4", "Col11a1", "Dpp4", "Fat4", "Aldh1a2")

# Assign a full gene name to the short geneID
genes <- data.frame(gene=rownames(seuratFB)) %>% mutate(geneID=gsub(".*\\.", "", gene)) %>% 
  filter(geneID %in% as.character(selMarker)) %>% arrange(factor(geneID, levels = rev(selMarker)))

# Scale the gene across all clusters
DotPlot(
  object        = seuratFB,
  features      = genes$gene,
  cols          = c("#2C3E50", "#F1E6A1"),
  col.min       = -2.5,
  col.max       =  2.5,
  dot.min       = 0,
  dot.scale     = 6,
  cluster.idents = TRUE,
  scale         = TRUE,
  scale.by      = "radius"
) +
  coord_flip() + scale_x_discrete(breaks = genes$gene,labels = genes$geneID) + scale_y_discrete(limits = ordVec) + theme(axis.text.x = element_text(angle = 90))

```
## Cluster abundances  {.tabset}

### Phenotypes
```{r, fig.height=3, fig.width=4}
# Plot abundance of clusters across Phenotypes
datList <- list()
for(con in unique(seuratFB$Phenotype)){
  seuratFBSub <- subset(seuratFB, Phenotype==con)
  dat_Phenotype <- as.data.frame(table(seuratFBSub$FBclusterLabel)) %>%
  mutate(percent=Freq/ncol(seuratFBSub)) %>% mutate(Phenotype=con)
  datList[[con]] <- dat_Phenotype
}
dat_all <- do.call("rbind", datList)

# Set Phenotype order
ordX <- c("healthy", "inflamed", "iCMP")

# Set cluster order
ordY <- c("Ccl19 FRCs", "Cxcl13 CoverCells", "Dpp4 Fibroblasts", "Ccl11 Fibroblasts", "Fat4 Fibroblasts", "Bmp4 Fibroblasts", "Tcf21 Fibroblasts", "Col11 Fibroblasts")
dat_all$Var1 <- factor(dat_all$Var1, levels = ordY)

# Barplot with Phenotypes on x axis and cluster colors
ggbarplot(dat_all, x= "Phenotype", y= "percent", fill = "Var1", legend = "right", legend.titel = "cluster", ylab = "frequency") +  scale_fill_manual(values = colPal_FB) + theme(axis.text.x = element_text(angle = 90)) + scale_x_discrete(limits=ordX) 
```

### Conditions
```{r, , fig.height=3, fig.width=4}
# Plot abundance of clusters across Conditions
datList <- list()
for(con in unique(seuratFB$cond)){
  seuratFBSub <- subset(seuratFB, cond==con)
  dat_con <- as.data.frame(table(seuratFBSub$FBclusterLabel)) %>%
  mutate(percent=Freq/ncol(seuratFBSub)) %>% mutate(cond=con)
  datList[[con]] <- dat_con
}
dat_all <- do.call("rbind", datList)

# Set cluster order
ordY <- c("Ccl19 FRCs", "Cxcl13 CoverCells", "Dpp4 Fibroblasts", "Ccl11 Fibroblasts", "Fat4 Fibroblasts", "Bmp4 Fibroblasts", "Tcf21 Fibroblasts", "Col11 Fibroblasts")
dat_all$Var1 <- factor(dat_all$Var1, levels = ordY)

# Barplot with Conditions on x axis and cluster colors
ggbarplot(dat_all, x= "cond", y= "percent", fill = "Var1", legend = "right", legend.titel = "cluster", ylab = "frequency")+ scale_fill_manual(values = colPal_FB)
```


# Calculate DE genes for each cluster
```{r, include = T, eval = F}
# Calculate DE genes per cluster compared to all others
Idents(seuratFB) <- seuratFB$FBclusterLabel
levels(seuratFB)

seuratFB_DEGenes <- FindAllMarkers(seuratFB, only.pos=T) %>% 
  dplyr::filter(p_val_adj < 0.01) 

# Save DEGenes as text file
basedir<- here()
basedir <- paste0(basedir, "/docs/")
write.table(
  seuratFB_DEGenes,
  file = paste0(basedir, "pericarditis_allmerged_FB_0.25_cL_DEGenes.txt"),
  row.names = FALSE,
  quote = FALSE,
  sep = "\t"
)
```

# Load DEGene text file
```{r}
basedir<- here()
fileNam <- paste0(basedir, "/docs/pericarditis_allmerged_FB_0.25_cL_DEGenes.txt")
seuratFB_DEGenes <- read.table(fileNam, header = TRUE, sep = "\t", stringsAsFactors = FALSE)
```


## Lollipop plot  {.tabset}

### Immune interaction pathways
```{r, fig.width=8}

# Clean DEGenes
DEGenes_all <- seuratFB_DEGenes %>% 
  mutate(EnsID = sub("\\..*", "", gene))

# Run enrichGO per cluster
cluster_list <- split(DEGenes_all, DEGenes_all$cluster)

ego_list <- lapply(names(cluster_list), function(clust){

  df <- cluster_list[[clust]] %>% 
    filter(avg_log2FC > 0)

  enrichGO(
    gene          = unique(df$EnsID),
    OrgDb         = org.Mm.eg.db,
    keyType       = "ENSEMBL",
    ont           = "BP",
    pAdjustMethod = "BH",
    pvalueCutoff  = 0.05
  )
})

names(ego_list) <- names(cluster_list)


# Selected immune interaction pathways
sel_all <- c(
  "cytokine-mediated signaling pathway",
  "regulation of innate immune response",
  "immune response-activating signaling pathway",
  "leukocyte cell-cell adhesion",
  "positive regulation of lymphocyte activation",
  "regulation of T cell proliferation"
)

# Combine to df
plot_df <- do.call(rbind, lapply(names(ego_list), function(cl){

  ego_all <- ego_list[[cl]]
  if (is.null(ego_all) || nrow(ego_all@result) == 0) return(NULL)

  ego_all@result %>%
    filter(Description %in% sel_all) %>%
    mutate(
      Cluster = cl,
      log10_adjP = -log10(p.adjust)
    )
}))


# Order clusters for plotting
cluster_colors <- c(
   "Cxcl13 CoverCells"="#CD6839",
   "Ccl19 FRCs"="#B2CD5B",
   "Col11 Fibroblasts"="#DCA86C",
   "Dpp4 Fibroblasts" = "#6B7F99",
   "Fat4 Fibroblasts"="#C1CDC1",
   "Bmp4 Fibroblasts"="#A2B5CD",
   "Ccl11 Fibroblasts"="#EEE8AA",
   "Tcf21 Fibroblasts"="#556B2F"
)
plot_df$Cluster <- factor(plot_df$Cluster, levels = names(cluster_colors))

# Lollipop plot
ggplot(plot_df,
       aes(x = log10_adjP,
           y = Description,
           color = Cluster)) +

  geom_linerange(
    aes(xmin = 0, xmax = log10_adjP),
    position = position_dodge(width = 0.8),
    linewidth = 0.8
  ) +

  geom_point(
    aes(size = log10_adjP),
    position = position_dodge(width = 0.8)
  ) +

  scale_color_manual(values = cluster_colors) +

  labs(
    x = expression(-log[10]("adj. p-value")),
    y = NULL,
    color = "Cluster",
    size = expression(-log[10]("adj. p-value"))
  ) +

  theme_minimal(base_size = 13) +
  theme(
    panel.grid.major.y = element_blank(),
    panel.grid.minor.y = element_blank()
  )
```

### ECM/Homeostasis pathways
```{r, fig.width=9}

# Clean DEGenes
DEGenes_all <- seuratFB_DEGenes %>% 
  mutate(EnsID = sub("\\..*", "", gene))

# Run enrichGO per cluster
cluster_list <- split(DEGenes_all, DEGenes_all$cluster)

ego_list <- lapply(names(cluster_list), function(clust){

  df <- cluster_list[[clust]] %>% 
    filter(avg_log2FC > 0)

  enrichGO(
    gene          = unique(df$EnsID),
    OrgDb         = org.Mm.eg.db,
    keyType       = "ENSEMBL",
    ont           = "BP",
    pAdjustMethod = "BH",
    pvalueCutoff  = 0.05
  )
})

names(ego_list) <- names(cluster_list)


# Selected niche homeostasis pathways
sel_all <- c(
  "extracellular matrix organization",
  "regulation of cellular response to growth factor stimulus",
  "transforming growth factor beta receptor signaling pathway",
  "regulation of angiogenesis",
  "connective tissue development",
  "ossification"
)

# Combine to df
plot_df <- do.call(rbind, lapply(names(ego_list), function(cl){

  ego_all <- ego_list[[cl]]
  if (is.null(ego_all) || nrow(ego_all@result) == 0) return(NULL)

  ego_all@result %>%
    filter(Description %in% sel_all) %>%
    mutate(
      Cluster = cl,
      log10_adjP = -log10(p.adjust)
    )
}))


# Order clusters for plotting
cluster_colors <- c(
   "Cxcl13 CoverCells"="#CD6839",
   "Ccl19 FRCs"="#B2CD5B",
   "Col11 Fibroblasts"="#DCA86C",
   "Dpp4 Fibroblasts" = "#6B7F99",
   "Fat4 Fibroblasts"="#C1CDC1",
   "Bmp4 Fibroblasts"="#A2B5CD",
   "Ccl11 Fibroblasts"="#EEE8AA",
   "Tcf21 Fibroblasts"="#556B2F"
)
plot_df$Cluster <- factor(plot_df$Cluster, levels = names(cluster_colors))

# Lollipop plot
ggplot(plot_df,
       aes(x = log10_adjP,
           y = Description,
           color = Cluster)) +

  geom_linerange(
    aes(xmin = 0, xmax = log10_adjP),
    position = position_dodge(width = 0.8),
    linewidth = 0.8
  ) +

  geom_point(
    aes(size = log10_adjP),
    position = position_dodge(width = 0.8)
  ) +

  scale_color_manual(values = cluster_colors) +

  labs(
    x = expression(-log[10]("adj. p-value")),
    y = NULL,
    color = "Cluster",
    size = expression(-log[10]("adj. p-value"))
  ) +

  theme_minimal(base_size = 13) +
  theme(
    panel.grid.major.y = element_blank(),
    panel.grid.minor.y = element_blank()
  )
```



# Save seuratFB object
```{r}
# Save seurat object
basedir <- here()
saveRDS(seuratFB, file = paste0(basedir, "/data/pericarditis_allmerged_FB_cleaned_cL_seurat.rds"))
```


# Session info
```{r}
sessionInfo()
date()
```


