---

title: "7_characterize FBs"
author: "Nadine"
date: "2025-04-09"
output: 
html_document:
  self_contained: no
  code_folding: show
  number_sections: no
  toc: yes
  toc_depth: 3
  toc_float: true
  collapsed: no
  smooth_scroll: yes
editor_options:
  chunk_output_type: inline
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(warning = FALSE, message = FALSE, dev = c("png", "pdf"))
seed <- 1234
```

#Load packages
```{r}
suppressPackageStartupMessages({
  library(ExploreSCdataSeurat3)
  library(runSeurat3)
  library(Seurat)
  library(ggpubr)
  library(pheatmap)
  library(SingleCellExperiment)
  library(dplyr)
  library(tidyverse)
  library(viridis)
  library(muscat)
  library(circlize)
  library(destiny)
  library(scater)
  library(metap)
  library(multtest)
  library(clusterProfiler)
  library(org.Mm.eg.db)
  library(msigdbr)
  library(enrichplot)
  library(DOSE)
  library(grid)
  library(gridExtra)
  library(ggupset)
  library(VennDiagram)
  library(reshape2)
  library(ggplot2)
  library(cowplot)
  library(purrr)
  library(here)
  library(ggsci)
  library(RColorBrewer)
  library(ggrepel)
})
```


#Load seurat object
```{r}
#Load merged and cleaned object containing FBs
basedir<- here()
fileNam <- paste0(basedir, "/data/pericarditis_allmerged_FB_cleaned_seurat.rds")
seuratFB <- readRDS(fileNam)
```

#Visualizaion in umaps
```{r}
Idents(seuratFB) <-  seuratFB$RNA_snn_res.0.2
DimPlot(seuratFB, reduction = "umap", label = T)

Idents(seuratFB) <-  seuratFB$RNA_snn_res.0.25
DimPlot(seuratFB, reduction = "umap", label = T)

```

#Marker genes per cluster
```{r, include = T, eval = T}
#Set identity to resolution 0.25
Idents(seuratFB) <-  seuratFB$RNA_snn_res.0.25

#Run Marker Genes
seuratFB_markers_all <- FindAllMarkers(object = seuratFB, assay ="RNA",
                                     only.pos = TRUE, min.pct = 0.25,
                                     logfc.threshold = 0.25,
                                     test.use = "wilcox")

#Save Marker Genes as Text File
basedir<- here()
basedir <- paste0(basedir, "/docs/")
write.table(
  seuratFB_markers_all,
  file = paste0(basedir, "pericarditis_allmerged_FB_cleaned_0.25_markerGenes.txt"),
  row.names = FALSE,
  quote = FALSE,
  sep = "\t"
)

```


#Assign labels to fibroblast clusters
```{r}
seuratFB$cluster <- seuratFB$RNA_snn_res.0.25

seuratFB$FBclusterLabel <- "FBclusterLabel"
seuratFB$FBclusterLabel[which(seuratFB$RNA_snn_res.0.25 %in% c("0"))] <- "Fat4 Fibroblasts"
seuratFB$FBclusterLabel[which(seuratFB$RNA_snn_res.0.25 %in% c("1"))] <- "Ccl19 FRCs"
seuratFB$FBclusterLabel[which(seuratFB$RNA_snn_res.0.25 %in% c("2"))] <- "Tcf21 Fibroblasts"
seuratFB$FBclusterLabel[which(seuratFB$RNA_snn_res.0.25 %in% c("3"))] <- "Cxcl13 CoverCells"
seuratFB$FBclusterLabel[which(seuratFB$RNA_snn_res.0.25 %in% c("4"))] <- "Ccl11 Fibroblasts"
seuratFB$FBclusterLabel[which(seuratFB$RNA_snn_res.0.25 %in% c("5"))] <- "Dpp4 Fibroblasts"
seuratFB$FBclusterLabel[which(seuratFB$RNA_snn_res.0.25 %in% c("6"))] <- "Bmp4 Fibroblasts"
seuratFB$FBclusterLabel[which(seuratFB$RNA_snn_res.0.25 %in% c("7"))] <- "Col11 Fibroblasts"


```


#Define color palette
```{r}
colPal_FB <- c("#C1CDC1", "#EEE8AA", "#556B2F", "#CD6839", "#B2CD5B", "#A2B5CD","#DCA86C", "#6B7F99")
names(colPal_FB) <- c("Fat4 Fibroblasts","Ccl11 Fibroblasts", "Tcf21 Fibroblasts", "Cxcl13 CoverCells", "Ccl19 FRCs", "Bmp4 Fibroblasts", "Col11 Fibroblasts", "Dpp4 Fibroblasts")

colPal_Phenotype <- c("#8B0000", "#FCDACB", "#E6E6E6")
names(colPal_Phenotype) <- c( "iCMP", "inflamed", "healthy")

```

#Stats
```{r}
table(seuratFB$orig.ident)
table(seuratFB$FBclusterLabel)
table(seuratFB$FBclusterLabel, seuratFB$cond)
table(seuratFB$Phenotype)
table(seuratFB$cond)
table(seuratFB$FBclusterLabel, seuratFB$SampleName)
```


#DimPlot FBclusterLabel
```{r}
Idents(seuratFB) <-  seuratFB$FBclusterLabel

DimPlot(seuratFB, reduction = "umap", cols=colPal_FB, group.by= "FBclusterLabel")
DimPlot(seuratFB, reduction = "umap", cols=colPal_FB, label = F, group.by= "FBclusterLabel", split.by = "cond")
```


#DimPlot Condition
```{r}
DimPlot(seuratFB, reduction = "umap", cols=c("#D3D3D3", "#B45B5C"), group.by = "cond")

DimPlot(seuratFB, reduction = "umap", cols=c("#D3D3D3", "#B45B5C"), group.by = "cond", split.by = "cond")

```

#DimPlot Phenotype
```{r}
DimPlot(seuratFB, reduction = "umap", cols=colPal_Phenotype, group.by = "Phenotype")

DimPlot(seuratFB, reduction = "umap", group.by = "Phenotype", cols = colPal_Phenotype, order = c("iCMP", "inflamed","healthy"), pt.size = 0.02)


DimPlot(seuratFB, reduction = "umap", cols=c("#E6E6E6","#E6E6E6","#FCDACB"), group.by = "Phenotype", order = "inflamed")
DimPlot(seuratFB, reduction = "umap", cols=c("#E6E6E6","#E6E6E6","#8B0000"), group.by = "Phenotype", order = "iCMP")

```


#DimPlot Age
```{r}
DimPlot(seuratFB, reduction = "umap", cols=c("#99B898", "#628395"), group.by = "age", split.by = "age")
DimPlot(seuratFB, reduction = "umap", cols=c("#D3D3D3", "#628395"), group.by = "age", order = "8wk")
DimPlot(seuratFB, reduction = "umap", cols=c("#D3D3D3", "#99B898"), group.by = "age", order = "4wk")
DimPlot(seuratFB, reduction = "umap", cols=c("#D3D3D3", "#B45B5C"), group.by = "cond", split.by = "age")

```

#Calculate marker genes per cluster
```{r, include = T, eval = T}
#Set identity to resolution FBclusterLabel
Idents(seuratFB) <-  seuratFB$FBclusterLabel

#Run Marker Genes
seuratFB_markers_all <- FindAllMarkers(object = seuratFB, assay ="RNA",
                                     only.pos = TRUE, min.pct = 0.25,
                                     logfc.threshold = 0.25,
                                     test.use = "wilcox")

#Save Marker Genes as Text file
basedir<- here()
basedir <- paste0(basedir, "/docs/")
write.table(
  seuratFB_markers_all,
  file = paste0(basedir, "pericarditis_allmerged_FB_0.25_cL_markerGenes.txt"),
  row.names = FALSE,
  quote = FALSE,
  sep = "\t"
)

```


#Load MarkerGenes
```{r}

basedir<- here()
fileNam <- paste0(basedir, "/docs/pericarditis_allmerged_FB_0.25_cL_markerGenes.txt")
seuratFB_MarkerGenes <- read.table(fileNam, header = TRUE, sep = "\t", stringsAsFactors = FALSE)
  
```

#Load average heatmap function
```{r avg heatmap funct}

avgHeatmap <- function(seuratFB, selGenes, colVecIdent, colVecCond=NULL,
                       ordVec=NULL, gapVecR=NULL, gapVecC=NULL,cc=FALSE,
                       cr=FALSE, condCol=FALSE){
  
  selGenes <- selGenes$gene
  
  ## assay data
  clusterAssigned <- as.data.frame(Idents(seuratFB)) %>%
  dplyr::mutate(cell=rownames(.))
  colnames(clusterAssigned)[1] <- "ident"
  seuratFBDat <- GetAssayData(seuratFB)
  
  ## genes of interest
  genes <- data.frame(gene=rownames(seuratFB)) %>% 
    mutate(geneID=gsub("^.*\\.", "", gene)) %>% filter(geneID %in% selGenes)

  ## matrix with averaged cnts per ident
  logNormExpres <- as.data.frame(t(as.matrix(
    seuratFBDat[which(rownames(seuratFBDat) %in% genes$gene),])))
  logNormExpres <- logNormExpres %>% dplyr::mutate(cell=rownames(.)) %>%
    dplyr::left_join(.,clusterAssigned, by=c("cell")) %>%
    dplyr::select(-cell) %>% dplyr::group_by(ident) %>%
    dplyr::summarise_all(mean)
  logNormExpresMa <- logNormExpres %>% dplyr::select(-ident) %>% as.matrix()
  rownames(logNormExpresMa) <- logNormExpres$ident
  logNormExpresMa <- t(logNormExpresMa)
  rownames(logNormExpresMa) <- gsub("^.*?\\.","",rownames(logNormExpresMa))
  
  ## remove genes if they are all the same in all groups
  ind <- apply(logNormExpresMa, 1, sd) == 0
  logNormExpresMa <- logNormExpresMa[!ind,]
  genes <- genes[!ind,]

  ## color columns according to cluster
  annotation_col <- as.data.frame(gsub("(^.*?_)","",
                                       colnames(logNormExpresMa)))%>%
    dplyr::mutate(celltype=gsub("(_.*$)","",colnames(logNormExpresMa)))
  colnames(annotation_col)[1] <- "col1"
  annotation_col <- annotation_col %>%
    dplyr::mutate(cond = gsub("(^[0-9]_?)","",col1)) %>%
    dplyr::select(cond, celltype)
  rownames(annotation_col) <- colnames(logNormExpresMa) 

  ann_colors = list(
      #cond = colVecCond,
      celltype=colVecIdent)
  if(is.null(ann_colors$cond)){
    annotation_col$cond <- NULL
  }
  
  ## adjust order
  logNormExpresMa <- logNormExpresMa[selGenes,]
  if(is.null(ordVec)){
    ordVec <- levels(seuratFB)
  }
  logNormExpresMa <- logNormExpresMa[,ordVec]

  ## scaled row-wise
  pheatmap(logNormExpresMa, scale="row" ,treeheight_row = 0, cluster_rows = cr, 
         cluster_cols = cc,
         color = colorRampPalette(c("#2166AC", "#F7F7F7", "#B2182B"))(50),
         annotation_col = annotation_col, cellwidth=15, cellheight=10,
         annotation_colors = ann_colors, gaps_row = gapVecR, gaps_col = gapVecC)
}
```

#Heatmap of top 10 marker genes per cluster
```{r, fig.height=15}
cluster <- levels(seuratFB)
selGenesAll <- seuratFB_MarkerGenes %>% group_by(cluster) %>% 
  top_n(-10, p_val_adj) %>% 
  top_n(10, avg_log2FC)
names(selGenesAll) [names(selGenesAll) == "gene"] <- "Gene"
selGenesAll <- selGenesAll %>% mutate(gene=gsub("^.*\\.", "", Gene)) %>% filter(nchar(Gene)>1)

colPal_FB <- c("#C1CDC1", "#EEE8AA", "#556B2F", "#CD6839", "#B2CD5B", "#A2B5CD","#DCA86C", "#6B7F99")
names(colPal_FB) <- c("Fat4 Fibroblasts","Ccl11 Fibroblasts", "Tcf21 Fibroblasts", "Cxcl13 CoverCells", "Ccl19 FRCs", "Bmp4 Fibroblasts", "Col11 Fibroblasts", "Dpp4 Fibroblasts")

ordVec <- names(colPal_FB)

Idents(seuratFB) <- seuratFB$FBclusterLabel
pOut <- avgHeatmap(seurat = seuratFB, selGenes = selGenesAll,
                  colVecIdent = colPal_FB, 
                  ordVec=ordVec,
                  gapVecR=NULL, gapVecC=NULL,cc=FALSE,
                  cr=T, condCol=F)

```



#Heatmap of selected immune interaction genes
```{r, fig.height=4, fig.width=5}

genes <- data.frame(gene=rownames(seuratFB))

selMarker <- list(stroma = c("Ccl19","Il6","Ccl2", "Cxcl1","Cxcl13", "Icam1", "Vcam1", "Il33", "Ccl11"))
    
Idents(seuratFB) <- seuratFB$FBclusterLabel

selGenes <- data.frame(gene=unlist(selMarker)) %>%
  rownames_to_column(var="grp") %>% mutate(Grp=gsub("_.*", "", grp))
grpCnt <- selGenes %>% group_by(Grp) %>% summarise(cnt=n())
gapR <- data.frame(Grp=unique(selGenes$Grp)) %>% 
  left_join(.,grpCnt, by="Grp") %>% mutate(cumSum=cumsum(cnt)) 

colPal_FB <- c("#C1CDC1", "#EEE8AA", "#556B2F", "#CD6839", "#B2CD5B", "#A2B5CD","#DCA86C", "#6B7F99")
names(colPal_FB) <- c("Fat4 Fibroblasts","Ccl11 Fibroblasts", "Tcf21 Fibroblasts", "Cxcl13 CoverCells", "Ccl19 FRCs", "Bmp4 Fibroblasts", "Col11 Fibroblasts", "Dpp4 Fibroblasts")

#ordVec <- c("Cxcl13 CoverCells", "Ccl19 FRCs", "Col11 Fibroblasts", "Dpp4 Fibroblasts", "Fat4 Fibroblasts","Bmp4 Fibroblasts", "Ccl11 Fibroblasts", "Tcf21 Fibroblasts")

pOut <- avgHeatmap(seuratFB = seuratFB, selGenes = selGenes,
                  colVecIdent = colPal_FB, 
                  ordVec=ordVec,
                  gapVecR=NULL, gapVecC=NULL,cc=T,
                  cr=T, condCol=F)
```








#Heatmap of selected immune interaction genes (ordered clusters)
```{r, fig.height=4, fig.width=5}

genes <- data.frame(gene=rownames(seuratFB))

selMarker <- list(stroma = c("Col1a1", "Col15a1", "Col11a1", "Eln", "Col8a1", "Col6a1", "Tgfbi", "Dcn", "Ogn"))
    
Idents(seuratFB) <- seuratFB$FBclusterLabel

selGenes <- data.frame(gene=unlist(selMarker)) %>%
  rownames_to_column(var="grp") %>% mutate(Grp=gsub("_.*", "", grp))
grpCnt <- selGenes %>% group_by(Grp) %>% summarise(cnt=n())
gapR <- data.frame(Grp=unique(selGenes$Grp)) %>% 
  left_join(.,grpCnt, by="Grp") %>% mutate(cumSum=cumsum(cnt)) 

colPal_FB <- c("#C1CDC1", "#EEE8AA", "#556B2F", "#CD6839", "#B2CD5B", "#A2B5CD","#DCA86C", "#6B7F99")
names(colPal_FB) <- c("Fat4 Fibroblasts","Ccl11 Fibroblasts", "Tcf21 Fibroblasts", "Cxcl13 CoverCells", "Ccl19 FRCs", "Bmp4 Fibroblasts", "Col11 Fibroblasts", "Dpp4 Fibroblasts")

ordVec <- c("Cxcl13 CoverCells", "Ccl19 FRCs", "Col11 Fibroblasts", "Dpp4 Fibroblasts", "Fat4 Fibroblasts","Bmp4 Fibroblasts", "Ccl11 Fibroblasts", "Tcf21 Fibroblasts")


pOut <- avgHeatmap(seuratFB = seuratFB, selGenes = selGenes,
                  colVecIdent = colPal_FB, 
                  ordVec=ordVec,
                  gapVecR=NULL, gapVecC=NULL,cc=F,
                  cr=T, condCol=F)
```



#Dotplot of selected marker genes
```{r, fig.height=4, fig.width=5}

Idents(seuratFB) <- seuratFB$FBclusterLabel

#Marker genes as simple vector
selMarker <-  c("Pdgfra", "Pdpn", "Cxcl13", "Ccl19", "Ccl11",  "Tcf21",  "Bmp4", "Col11a1", "Dpp4", "Fat4")

#Assign a full gene name to the short geneID
genes <- data.frame(gene=rownames(seuratFB)) %>% mutate(geneID=gsub(".*\\.", "", gene)) %>% 
  filter(geneID %in% as.character(selMarker)) %>% arrange(factor(geneID, levels = rev(selMarker)))

#Create dataframe of cL
clusterDat <- data.frame(cluster=levels(seuratFB)) %>%
  mutate(label = levels(seuratFB))


#Scale the gene across all clusters
DotPlot(
  object=seuratFB,
  features = genes$gene,
  cols = c("#2C3E50", "#F1E6A1"),
  col.min = -2.5,
  col.max = 2.5,
  dot.min = 0,
  dot.scale = 6,
  idents = NULL,
  group.by = NULL,
  split.by = NULL,
  cluster.idents = T, #TRUE= automatic order, FALSE= ordered 
  scale = T,
  scale.by = "radius",
  scale.min = NA,
  scale.max = NA
) + coord_flip() + scale_x_discrete(breaks=genes$gene, labels=genes$geneID)  + scale_y_discrete(breaks=clusterDat$cluster, labels=clusterDat$label) + theme(axis.text.x = element_text(angle = 90))

```

#Dotplot of selected marker genes (ordered clusters)
```{r, fig.height=4, fig.width=5}

ordVec <- c("Cxcl13 CoverCells", "Ccl19 FRCs", "Col11 Fibroblasts", "Dpp4 Fibroblasts", "Fat4 Fibroblasts","Bmp4 Fibroblasts", "Ccl11 Fibroblasts", "Tcf21 Fibroblasts")


Idents(seuratFB) <- seuratFB$FBclusterLabel
Idents(seuratFB) <- factor(Idents(seuratFB), levels = ordVec)

#Marker genes as simple vector
selMarker <-  c("Pdgfra", "Pdpn", "Cxcl13", "Ccl19", "Ccl11",  "Tcf21",  "Bmp4", "Col11a1", "Dpp4", "Fat4")


#Assign a full gene name to the short geneID
genes <- data.frame(gene=rownames(seuratFB)) %>% mutate(geneID=gsub(".*\\.", "", gene)) %>% 
  filter(geneID %in% as.character(selMarker)) %>% arrange(factor(geneID, levels = rev(selMarker)))

#Scale the gene across all clusters
DotPlot(
  object        = seuratFB,
  features      = genes$gene,
  cols          = c("#2C3E50", "#F1E6A1"),
  col.min       = -2.5,
  col.max       =  2.5,
  dot.min       = 0,
  dot.scale     = 6,
  cluster.idents = TRUE,
  scale         = TRUE,
  scale.by      = "radius"
) +
  coord_flip() + scale_x_discrete(breaks = genes$gene,labels = genes$geneID) + scale_y_discrete(limits = ordVec) + theme(axis.text.x = element_text(angle = 90))

```

#Abundance across all samples
```{r}
#table(seuratFB$SampleName)

#Plot abundance of clusters in individual samples
datList <- NULL
for(con in unique(seuratFB$SampleName)){
  seuratFBSub <- subset(seuratFB, SampleName==con)
  print(dim(seuratFBSub))
  dat_SampleName <- as.data.frame(table(seuratFBSub$FBclusterLabel)) %>%
  mutate(percent=Freq/ncol(seuratFBSub)) %>% mutate(SampleName=con)
  datList[[con]] <- dat_SampleName
}
dat_all <- do.call("rbind", datList)


ordX <- c("4w_LMC1", "4w_LMC2", "4w_LMC3", "4w_LMC4", "8w_LMC1", "8w_LMC2", "8w_LMC3", "8w_LMC4", "8w_LMC5", "8w_LMC6", "8w_LMC7", "8w_LMC8" , "4w_TCRM1", "4w_TCRM2", "4w_TCRM3", "4w_TCRM4", "4w_TCRM5", "4w_TCRM6", "4w_TCRM7", "8w_TCRM1", "8w_TCRM2", "8w_TCRM3", "8w_TCRM4", "8w_TCRM5", "8w_TCRM6", "8w_TCRM7", "8w_TCRM8")

#Create barplot with samples on x axis and cluster colors
ggbarplot(dat_all, x= "SampleName", y= "percent", fill = "Var1", legend = "right", legend.titel = "cluster", ylab = "frequency") +  scale_fill_manual(values = colPal_FB) + theme(axis.text.x = element_text(angle = 90)) + scale_x_discrete(limits=ordX) 



```

#Abundance across non-sorted samples
```{r}
#Plot abundance of clusters in individual samples
##Leave out sorted samples

seuratFBSub <- subset(seuratFB, SampleName %in% c("8w_LMC8", "8w_TCRM7", "8w_TCRM8"), invert=TRUE )
  
##run calculations
datList <- NULL
for(con in unique(seuratFBSub$SampleName)){
  seuratFB_sub <- subset(seuratFBSub, SampleName==con)
  print(dim(seuratFB_sub))
  dat_SampleName <- as.data.frame(table(seuratFB_sub$FBclusterLabel)) %>%
  mutate(percent=Freq/ncol(seuratFB_sub)) %>% mutate(SampleName=con)
  datList[[con]] <- dat_SampleName
}
dat_all <- do.call("rbind", datList)


ordX <- c("4w_LMC1", "4w_LMC2", "4w_LMC3", "4w_LMC4", "8w_LMC1", "8w_LMC2", "8w_LMC3", "8w_LMC4", "8w_LMC5", "8w_LMC6", "8w_LMC7",  "4w_TCRM1", "4w_TCRM2", "4w_TCRM3", "4w_TCRM4", "4w_TCRM5", "4w_TCRM6", "4w_TCRM7", "8w_TCRM1", "8w_TCRM2", "8w_TCRM3", "8w_TCRM4", "8w_TCRM5", "8w_TCRM6")

#Create barplot with samples on x axis and cluster colors
ggbarplot(dat_all, x= "SampleName", y= "percent", fill = "Var1", legend = "right", legend.titel = "cluster", ylab = "frequency") +  scale_fill_manual(values = colPal_FB) + theme(axis.text.x = element_text(angle = 90)) + scale_x_discrete(limits=ordX) 



```

#Abundance across phenotypes
```{r, fig.height=3, fig.width=4}

datList <- NULL
for(con in unique(seuratFB$Phenotype)){
  seuratFBSub <- subset(seuratFB, Phenotype==con)
  print(dim(seuratFBSub))
  dat_Phenotype <- as.data.frame(table(seuratFBSub$FBclusterLabel)) %>%
  mutate(percent=Freq/ncol(seuratFBSub)) %>% mutate(Phenotype=con)
  datList[[con]] <- dat_Phenotype
}
dat_all <- do.call("rbind", datList)

#Set order of Phenotypes
ordX <- c("healthy", "inflamed", "iCMP")

#Create barplot with Phenotypes on x axis and cluster colors
ggbarplot(dat_all, x= "Phenotype", y= "percent", fill = "Var1", legend = "right", legend.titel = "cluster", ylab = "frequency") +  scale_fill_manual(values = colPal_FB) + theme(axis.text.x = element_text(angle = 90)) + scale_x_discrete(limits=ordX) 
```

#Abundance across conditions
```{r, , fig.height=3, fig.width=4}
datList <- NULL
for(con in unique(seuratFB$cond)){
  seuratFBSub <- subset(seuratFB, cond==con)
  print(dim(seuratFBSub))
  dat_con <- as.data.frame(table(seuratFBSub$FBclusterLabel)) %>%
  mutate(percent=Freq/ncol(seuratFBSub)) %>% mutate(cond=con)
  datList[[con]] <- dat_con
}
dat_all <- do.call("rbind", datList)

## plot abundance
ggbarplot(dat_all, x= "cond", y= "percent", fill = "Var1", legend = "right", legend.titel = "cluster", ylab = "frequency")+ scale_fill_manual(values = colPal_FB)

```

#Calculate DE genes for each cluster
```{r, include = T, eval = F}
## Calculate DE genes per cluster compared to all others
Idents(seuratFB) <- seuratFB$FBclusterLabel
levels(seuratFB)

seuratFB_DEGenes <- FindAllMarkers(seuratFB, only.pos=T) %>% 
  dplyr::filter(p_val_adj < 0.01) 

#Save DEGenes as Text file
basedir<- here()
basedir <- paste0(basedir, "/docs/")
write.table(
  seuratFB_DEGenes,
  file = paste0(basedir, "pericarditis_allmerged_FB_0.25_cL_DEGenes.txt"),
  row.names = FALSE,
  quote = FALSE,
  sep = "\t"
)

```

#Load DEGenes
```{r}
basedir<- here()
fileNam <- paste0(basedir, "/docs/pericarditis_allmerged_FB_0.25_cL_DEGenes.txt")
seuratFB_DEGenes <- read.table(fileNam, header = TRUE, sep = "\t", stringsAsFactors = FALSE)

```

#Lollipop plot - immune interaction
```{r, fig.width=8}

#Clean DEGenes
DEGenes_all <- seuratFB_DEGenes %>% 
  mutate(EnsID = sub("\\..*", "", gene))

#Run enrichGO per cluster
cluster_list <- split(DEGenes_all, DEGenes_all$cluster)

ego_list <- lapply(names(cluster_list), function(clust){

  df <- cluster_list[[clust]] %>% 
    filter(avg_log2FC > 0)

  enrichGO(
    gene          = unique(df$EnsID),
    OrgDb         = org.Mm.eg.db,
    keyType       = "ENSEMBL",
    ont           = "BP",
    pAdjustMethod = "BH",
    pvalueCutoff  = 0.05
  )
})

names(ego_list) <- names(cluster_list)


#Selected immune interaction pathways
sel_all <- c(
  "cytokine-mediated signaling pathway",
  "regulation of innate immune response",
  "immune response-activating signaling pathway",
  "leukocyte cell-cell adhesion",
  "positive regulation of lymphocyte activation",
  "regulation of T cell proliferation"
)

#Combine to df
plot_df <- do.call(rbind, lapply(names(ego_list), function(cl){

  ego_all <- ego_list[[cl]]
  if (is.null(ego_all) || nrow(ego_all@result) == 0) return(NULL)

  ego_all@result %>%
    filter(Description %in% sel_all) %>%
    mutate(
      Cluster = cl,
      log10_adjP = -log10(p.adjust)
    )
}))


#Order clusters for plotting
cluster_colors <- c(
   "Cxcl13 CoverCells"="#CD6839",
   "Ccl19 FRCs"="#B2CD5B",
   "Col11 Fibroblasts"="#DCA86C",
   "Dpp4 Fibroblasts" = "#6B7F99",
   "Fat4 Fibroblasts"="#C1CDC1",
   "Bmp4 Fibroblasts"="#A2B5CD",
   "Ccl11 Fibroblasts"="#EEE8AA",
   "Tcf21 Fibroblasts"="#556B2F"
)
plot_df$Cluster <- factor(plot_df$Cluster, levels = names(cluster_colors))

#Lollipop plot
ggplot(plot_df,
       aes(x = log10_adjP,
           y = Description,
           color = Cluster)) +

  geom_linerange(
    aes(xmin = 0, xmax = log10_adjP),
    position = position_dodge(width = 0.8),
    linewidth = 0.8
  ) +

  geom_point(
    aes(size = log10_adjP),
    position = position_dodge(width = 0.8)
  ) +

  scale_color_manual(values = cluster_colors) +

  labs(
    x = expression(-log[10]("adj. p-value")),
    y = NULL,
    color = "Cluster",
    size = expression(-log[10]("adj. p-value"))
  ) +

  theme_minimal(base_size = 13) +
  theme(
    panel.grid.major.y = element_blank(),
    panel.grid.minor.y = element_blank()
  )


```


#Lollipop plot - ECM/Homeostasis
```{r, fig.width=9}

#Clean DEGenes
DEGenes_all <- seuratFB_DEGenes %>% 
  mutate(EnsID = sub("\\..*", "", gene))

#Run enrichGO per cluster
cluster_list <- split(DEGenes_all, DEGenes_all$cluster)

ego_list <- lapply(names(cluster_list), function(clust){

  df <- cluster_list[[clust]] %>% 
    filter(avg_log2FC > 0)

  enrichGO(
    gene          = unique(df$EnsID),
    OrgDb         = org.Mm.eg.db,
    keyType       = "ENSEMBL",
    ont           = "BP",
    pAdjustMethod = "BH",
    pvalueCutoff  = 0.05
  )
})

names(ego_list) <- names(cluster_list)


#Selected niche homeostasis pathways
sel_all <- c(
  "extracellular matrix organization",
  "regulation of cellular response to growth factor stimulus",
  "transforming growth factor beta receptor signaling pathway",
  "regulation of angiogenesis",
  "connective tissue development",
  "ossification"
)

#Combine to df
plot_df <- do.call(rbind, lapply(names(ego_list), function(cl){

  ego_all <- ego_list[[cl]]
  if (is.null(ego_all) || nrow(ego_all@result) == 0) return(NULL)

  ego_all@result %>%
    filter(Description %in% sel_all) %>%
    mutate(
      Cluster = cl,
      log10_adjP = -log10(p.adjust)
    )
}))


#Order clusters for plotting
cluster_colors <- c(
   "Cxcl13 CoverCells"="#CD6839",
   "Ccl19 FRCs"="#B2CD5B",
   "Col11 Fibroblasts"="#DCA86C",
   "Dpp4 Fibroblasts" = "#6B7F99",
   "Fat4 Fibroblasts"="#C1CDC1",
   "Bmp4 Fibroblasts"="#A2B5CD",
   "Ccl11 Fibroblasts"="#EEE8AA",
   "Tcf21 Fibroblasts"="#556B2F"
)
plot_df$Cluster <- factor(plot_df$Cluster, levels = names(cluster_colors))

#Lollipop plot
ggplot(plot_df,
       aes(x = log10_adjP,
           y = Description,
           color = Cluster)) +

  geom_linerange(
    aes(xmin = 0, xmax = log10_adjP),
    position = position_dodge(width = 0.8),
    linewidth = 0.8
  ) +

  geom_point(
    aes(size = log10_adjP),
    position = position_dodge(width = 0.8)
  ) +

  scale_color_manual(values = cluster_colors) +

  labs(
    x = expression(-log[10]("adj. p-value")),
    y = NULL,
    color = "Cluster",
    size = expression(-log[10]("adj. p-value"))
  ) +

  theme_minimal(base_size = 13) +
  theme(
    panel.grid.major.y = element_blank(),
    panel.grid.minor.y = element_blank()
  )


```



#Save seuratFB object
```{r}
#Save seurat object
basedir <- here()
saveRDS(seuratFB, file = paste0(basedir, "/data/pericarditis_allmerged_FB_cleaned_cL_seurat.rds"))
```


#Session info
```{r}
sessionInfo()
date()
```


