---
title: "8_FRC Ctrl vs TCRM"
author: "Nadine"
date: "2025-04-09"
output: 
html_document:
  self_contained: no
  code_folding: show
  number_sections: no
  toc: yes
  toc_depth: 3
  toc_float: true
  collapsed: no
  smooth_scroll: yes
editor_options:
  chunk_output_type: inline
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(warning = FALSE, message = FALSE, dev = c("png", "pdf"))
seed <- 1234
```

#Load packages
```{r}
suppressPackageStartupMessages({
  library(ExploreSCdataSeurat3)
  library(runSeurat3)
  library(Seurat)
  library(ggpubr)
  library(pheatmap)
  library(SingleCellExperiment)
  library(dplyr)
  library(tidyverse)
  library(viridis)
  library(muscat)
  library(circlize)
  library(destiny)
  library(scater)
  library(metap)
  library(multtest)
  library(clusterProfiler)
  library(org.Mm.eg.db)
  library(msigdbr)
  library(enrichplot)
  library(DOSE)
  library(grid)
  library(gridExtra)
  library(ggupset)
  library(VennDiagram)
  library(reshape2)
  library(ggplot2)
  library(cowplot)
  library(purrr)
  library(here)
  library(ggsci)
  library(RColorBrewer)
  library(ggrepel)
  library(ImmunoLogic)
})
```


#Load seurat object
```{r}
#Load merged and cleaned object containing FBs
basedir<- here()
fileNam <- paste0(basedir, "/data/pericarditis_allmerged_FB_cleaned_cL_seurat.rds")
seuratFB <- readRDS(fileNam)
```

#Define color palette
```{r}
colPal_FB <- c("#C1CDC1", "#EEE8AA", "#556B2F", "#CD6839", "#B2CD5B", "#A2B5CD","#DCA86C", "#6B7F99")
names(colPal_FB) <- c("Fat4 Fibroblasts","Ccl11 Fibroblasts", "Tcf21 Fibroblasts", "Cxcl13 CoverCells", "Ccl19 FRCs", "Bmp4 Fibroblasts", "Col11 Fibroblasts", "Dpp4 Fibroblasts")

colPal_Phenotype <- c("#8B0000", "#FCDACB", "#E6E6E6")
names(colPal_Phenotype) <- c( "iCMP", "inflamed", "healthy")

```


#calculate DE genes between conditions per cluster
```{r, include = T, eval = F}
#Set identity
Idents(seuratFB) <- seuratFB$FBclusterLabel

clusters <- levels(seuratFB)
conditions <- unique(seuratFB$cond)

#Initialize list to store results
DE_list <- list()

#Loop over each cluster
for (cl in clusters) {
  cat("Processing cluster:", cl, "\n")
  
  #Subset to the current cluster
  sub <- subset(seuratFB, idents = cl)
  
  #Set condition as active identity class for comparison
  Idents(sub) <- sub$cond
  
#Set identity as condition
  #DE genes between conditions within this cluster
  deg <- FindMarkers(sub,
                     ident.1 = "TCRM",
                     ident.2 = "LMC",
                     only.pos = FALSE,
                     logfc.threshold = 0,
                     min.pct = 0.1)
  
   deg <- deg %>% tibble::rownames_to_column(var = "gene")
  
  
  #Add cluster label
  deg$cluster <- cl
  deg$comparison <- paste("TCRM", "vs", "LMC")
  
  #Store in list
  DE_list[[cl]] <- deg
}

#Combine all results
DE_all <- bind_rows(DE_list, .id = "cluster_id")

#Filter for significant DE genes
DE_all_filtered <- DE_all %>% filter(p_val_adj < 0.01)


#Save DE genes as text file
basedir<- here()
basedir <- paste0(basedir, "/docs/")
write.table(
  DE_all_filtered,
  file = paste0(basedir, "pericarditis_allmerged_FB_cleaned_cL_DEgenes_perCluster_perCondition.txt"),
  row.names = FALSE,
  quote = FALSE,
  sep = "\t",
  col.names = TRUE
)


```

#Load DEGenes
```{r}

basedir<- here()
fileNam <- paste0(basedir, "/docs/pericarditis_allmerged_FB_cleaned_cL_DEgenes_perCluster_perCondition.txt")
DE_all_filtered <- read.table(fileNam, header = TRUE, sep = "\t", stringsAsFactors = FALSE)

  
```


#Barplot - number of DE genes per cluster
```{r}
DE_summary <- DE_all_filtered %>%
  group_by(cluster) %>%
  summarise(n_DE = n())

ggplot(DE_summary, aes(x = cluster, y = n_DE, fill = cluster)) +
  
  geom_bar(stat = "identity") +
  theme_classic() +
  labs(title = "Number of DEGs per cluster")+
  scale_fill_manual(values = colPal_FB)

```

#Transform gene names to short version
```{r}

DE_all_filtered <- DE_all_filtered %>%
  mutate(
    EnsID_clean = sub("\\..*$", "", gene),   # keep ENSMUSG ID only
    gene_clean  = sub("^[^.]+\\.", "", gene) # keep gene symbol only
  )


```


#GSEA Ccl19 FRCs
```{r, fig.height=8}
DEGenes_FRC <- DE_all_filtered %>%
  filter(cluster == "Ccl19 FRCs") %>%
  filter(!is.na(avg_log2FC)) %>%
  mutate(Gene = gene_clean)

gene_list <- DEGenes_FRC$avg_log2FC
names(gene_list) <- DEGenes_FRC$EnsID_clean
gene_list <- gene_list[!duplicated(names(gene_list))]
gene_list <- sort(gene_list, decreasing = TRUE)


ego_FRC_GSEA <- gseGO(
  geneList = gene_list,
  OrgDb = org.Mm.eg.db,
  keyType = "ENSEMBL",
  ont = "BP",
  pAdjustMethod = "BH",
  pvalueCutoff = 0.05,
  verbose = FALSE
)

#Convert ENSMUS IDs to gene symbols
ego_FRC_GSEA <- setReadable(ego_FRC_GSEA, OrgDb = org.Mm.eg.db, keyType = "ENSEMBL")

#Filter for significant results
ego_FRC_GSEA <- simplify(ego_FRC_GSEA, cutoff= 0.7, by = "p.adjust", select_fun = min)
ego_FRC_GSEA@result <- ego_FRC_GSEA@result [ego_FRC_GSEA@result$p.adjust < 0.05,]

dotplot(ego_FRC_GSEA, showCategory = 20) +
  ggtitle("GSEA: Ccl19 FRCs (TCRM vs LMC)")


```

#dotplot selected pathways FRCs
```{r}
#head(ego_FRC_GSEA@result$Description, 20)

#Filter selected pathways
sel_FRC <- c(
  "regulation of lymphocyte activation",
  "positive regulation of T cell activation",
  "positive regulation of immune response",
  "antigen processing and presentation",
  "positive regulation of leukocyte cell-cell adhesion",
  "MHC protein complex assembly",
  "response to interferon-beta"
)

ego_fil <- ego_FRC_GSEA@result %>%
  filter(Description %in% sel_FRC)

#Dotplot
##size = gene count, color = NES
dotplot(ego_FRC_GSEA, showCategory = sel_FRC, color = "p.adjust") +
  ggtitle("Selected GO: Ccl19 FRCs (TCRM vs. LMC)")



```


#Calculate DE genes per cluster and condition and split the list per condition
```{r, include = T, eval = F}

#Set identity
Idents(seuratFB) <- seuratFB$FBclusterLabel

clusters <- levels(seuratFB)
conditions <- unique(seuratFB$cond)

DEgenes_cond_list <- list()

for (cl in clusters) {
  cat("Processing cluster:", cl, "\n")
  
  #Subset to this cluster
  sub <- subset(seuratFB, idents = cl)
  
  #Set condition as identity class
  Idents(sub) <- sub$cond
  
  #Ensure both conditions exist in this cluster
  cond_present <- intersect(conditions, levels(Idents(sub)))
  if (length(cond_present) < 2) {
    cat("   Skipping — not both conditions present in", cl, "\n")
    next
  }
  
  #Differential expression: TCRM vs LMC
  deg <- FindMarkers(
    sub,
    ident.1 = "TCRM",
    ident.2 = "LMC",
    only.pos = FALSE,
    logfc.threshold = 0,
    min.pct = 0.1
  )
  
  #Add gene column and extract EnsID
  deg <- deg %>%
    tibble::rownames_to_column(var = "gene") %>%
    mutate(
      EnsID = sub("\\..*", "", gene),  # clean Ensembl ID
      Cluster = cl
    )
  
  #Split into condition-specific sets
  deg_up_TCRM <- deg %>% filter(avg_log2FC > 0) %>% mutate(Condition = "TCRM")
  deg_up_LMC  <- deg %>% filter(avg_log2FC < 0) %>% mutate(Condition = "LMC")
  
  #Store
  DEgenes_cond_list[[paste0(cl, "_TCRM")]] <- deg_up_TCRM
  DEgenes_cond_list[[paste0(cl, "_LMC")]]  <- deg_up_LMC
}

#Combine all results
DEgenes_cond_all <- bind_rows(DEgenes_cond_list)


write.table(DEgenes_cond_all,
            file = paste0(basedir, "/docs/pericarditis_allmerged_FB_cleaned_cL_DEgenes_perCluster_perCondition_individual.txt"),
            row.names = FALSE, quote = FALSE, sep = "\t", col.names = TRUE)

```

#Load DEGenes
```{r}

basedir<- here()
fileNam <- paste0(basedir, "/docs/pericarditis_allmerged_FB_cleaned_cL_DEgenes_perCluster_perCondition_individual.txt")
DEgenes_cond_all <- read.table(fileNam, header = TRUE, sep = "\t", stringsAsFactors = FALSE)


```

#GSEA for all clusters per condition
```{r}
cluster_cond_list <- split(DEgenes_cond_all, list(DEgenes_cond_all$Cluster, DEgenes_cond_all$Condition))

ego_list <- lapply(names(cluster_cond_list), function(nm){
  df <- cluster_cond_list[[nm]]
  if (nrow(df) < 5) return(NULL)
  
  enrichGO(
    gene = unique(df$EnsID),
    OrgDb = org.Mm.eg.db,
    keyType = "ENSEMBL",
    ont = "BP",
    pAdjustMethod = "BH",
    pvalueCutoff = 0.05
  )
})
names(ego_list) <- names(cluster_cond_list)

```


```{r}
#Select pathways
sel_FRC <- c(
  "regulation of lymphocyte activation",
  "positive regulation of T cell activation",
  "positive regulation of cell-cell adhesion",
  "positive regulation of immune response",
  "antigen processing and presentation",
  "antigen processing and presentation of exogenous antigen",
  "regulation of leukocyte cell-cell adhesion"
)

#Combine enrichment results across clusters and conditions
plot_df <- do.call(
  rbind,
  lapply(names(ego_list), function(nm) {

    ego <- ego_list[[nm]]
    if (is.null(ego)) return(NULL)

    df <- ego@result

    #Extract cluster and condition from list name (Cluster.Condition)
    parts <- strsplit(nm, "\\.")[[1]]
    cluster <- parts[1]
    condition <- parts[2]

    df %>%
      filter(Description %in% sel_FRC) %>%
      mutate(
        Cluster   = cluster,
        Condition = condition,
        log10_padj = -log10(p.adjust)
      )
  })
)

#Filter for Ccl19 FRCs and conditions
plot_df_cluster <- plot_df %>%
  filter(
    Cluster == "Ccl19 FRCs",
    Condition %in% c("LMC", "TCRM")
  )

#Filter
plot_df_cluster <- plot_df_cluster %>%
  mutate(log10_padj = pmin(log10_padj, 10))

#Order pathways by maximum significance across conditions
plot_df_cluster <- plot_df_cluster %>%
  group_by(Description) %>%
  mutate(max_log10_padj = max(log10_padj, na.rm = TRUE)) %>%
  ungroup() %>%
  arrange(desc(max_log10_padj)) %>%
  mutate(Description = factor(Description, levels = unique(Description)))

```


#Heatmap on selected pathways
```{r}
# Keep only Ccl19 FRCs
plot_df_cluster <- plot_df %>%
  filter(
    Cluster == "Ccl19 FRCs",
    Condition %in% c("LMC", "TCRM")
  ) %>%
 
  mutate(log10_padj = pmin(-log10(p.adjust), 10))

#Order pathways by maximum significance
plot_df_cluster <- plot_df_cluster %>%
  group_by(Description) %>%
  mutate(MaxSig = max(log10_padj, na.rm = TRUE)) %>%
  ungroup() %>%
  arrange(desc(MaxSig)) %>%
  mutate(Description = factor(Description, levels = unique(Description)))

#Plot heatmap
ggplot(
  plot_df_cluster,
  aes(x = Condition, y = Description, fill = log10_padj)
) +
  geom_tile(color = "#FFFFFF", linewidth = 0.3) +
  scale_fill_gradient(
    low = "#F2F2F2",
    high = "#B45B5C",
    na.value = "#FFFFFF",
    breaks = c(1.3, 2, 3, 5, 10),
    labels = c("0.05", "0.01", "0.001", "1e-5", "<1e-10")
  ) +
  labs(
    x = NULL,
    y = NULL,
    fill = expression(-log[10]("adj p")),
    title = "Ccl19 FRCs — LMC vs. TCRM pathway enrichment"
  ) +
  theme_minimal(base_size = 14) +
  theme(
    axis.text.x = element_text(size = 12, face = "bold"),
    axis.text.y = element_text(size = 11),
    panel.grid = element_blank(),
    plot.title = element_text(face = "bold", hjust = 0.5)
  )

```



#Session info
```{r}
sessionInfo()
date()
```


